### 2023-02-17 paper.qmd
First edits to past document.

---
title: "Writing a reproducible paper with Quarto"
author: "Paul C. Bauer, Camille Landesvatter"
format: html
date: 'First version: 16 February, 2023<br>This version: `r format(Sys.time(), "%d %B, %Y")`'
abstract: 'The present paper provides a template to write a reproducible scientific paper with R Markdown and Pagedown.^[Based on an earlier R Markdown template that uses Latex and can be downloaded under https://osf.io/q395s [see @Bauer2018-dl].] Below we outline some of the "tricks"/code (e.g., referencing tables, sections etc.) we had to figure out to produce this document. The underlying files which produce this document can be [downloaded here](https://drive.google.com/drive/folders/1rWzj-Bu1EKqkSuE1gaFzduBJhzlThpkw?usp=sharing). Importantly, we also provide different CSS and HTML files that can be used to achieve a pdf output with the look of a "working paper". We are convinced that in the future there will be many improvements and developments with regards to RStudio, R markdown and Pagedown. We intend to update this file when we discover more convenient code. You can follow any updates on the [github repository](https://github.com/paulcbauer/Writing_a_reproducable_paper_with_quarto/).'
bibliography: references.bib
tbl-cap-location: top
execute: 
  echo: false
  warning: false
  message: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = FALSE)
# Use chache = TRUE if you want to speed up compilation
# A function to allow for showing some of the inline code
rinline <- function(code){
  html <- '<code  class="r">``` `r CODE` ```</code>'
  sub("CODE", code, html)
}
```

# Why reproducible research (in R)?
Some arguments...

* **Access**: Research is normally funded by taxpayers (researchers are also taxpayers). Hence, it should be freely accessible to everyone without any barriers, e.g., without requiring commercial software. Importantly, researchers from developing countries are even more dependent on free access to knowledge [@Kirsop2005-ro].

* **Reproducability**: Even if you have written a study and analyzed the data yourself you will forget what you did after a few months. A fully reproducible setup will help you to trace back your own steps. Obviously, the same is true for other researchers who may want to understand your work and built on it. It may sound like a joke but why not aim for a document that can be used to reproduce your findings in 500 years.

* **Errors**: Manual steps in data analysis (e.g., manually copy/pasting values into a table etc.) may introduce errors. R Markdown allows you to **automatize** such steps and/or avoid them.

* **Revisions**: Revising a paper takes much less time if you have all the code you need in one place, i.e., one `.qmd` file. For instance, if you decide to exclude a subset of your data you simply need to insert one line of your code at the beginning and everything is rebuilt/re-estimated automatically.

# Why Quarto?

Formatting text as PDF is probably one of the most widespread standards in the scientific community, especially when it comes to submitting papers and similar documents. The traditional way to well-formatted and good-looking PDFs is often through LaTeX or Word. However, if you have spent hours and hours debugging latex code (or getting it to run) you may be on the lookout for something new.

The fairly new `pagedown` R package takes a completely new approach. While the main purpose of pagedown is to create high-quality PDFs, the idea is to take advantage of modern web technologies (HTML/JSS/Javascript) with which one can design web pages and eventually print those to PDF.

While web pages are usually single-page scrollable documents, pagedown uses the JavaScript library `Paged.js` which allows documents to be paginated with elements like headers, footers and everything a readable scientific paper will need. Additionally, pagedown documents are based on R Markdown. In our view, Pagedown and the underlying technology may replace Latex in the long run. In the near future it should also be possible to produce a PDF with static graph and an equivalent html with interactive graps (see dicsussion [here](https://github.com/rstudio/pagedown/pull/87)).

# Prerequesites

We assume that you are using R on a day-to-day basis and you may have even started to work in R Markdown. If you don't know what R Markdown is there are many great resources that you should use (e.g. watch this [short video](https://vimeo.com/178485416)). An older template [see @Bauer2018-dl; https://osf.io/q395s] on which this newer template is based, may provide a quick entry point to writing a reproducible with R Markdown and Latex.

Based on R Markdown, Pagedown allows you to create custom and well-formatted (paged) HTML Documents. For a comprehensive overview watch [this video](https://www.rstudio.com/resources/rstudioconf-2019/pagedown-creating-beautiful-pdfs-with-r-markdown-and-css/) which is a record of a talk introducing `pagedown` given by Yihui Xie (who in addition to Romain Lesur developed the `pagedown` package). If you are not in a video watching mood find the slides [here](https://slides.yihui.org/2019-rstudio-conf-pagedown.html#1).

Then...

* ...install [R](https://www.r-project.org/) and [Rstudio](https://www.rstudio.com/) (most recent versions) [@R2017; @Rstudio2015].
* ...install the "pagedown"-package from github using the code below [@Xie2021-bi; @Xie2021-ls].
```{r eval=FALSE, include=TRUE}
remotes::install_github('rstudio/pagedown')
```
* ...also install the packages below using the code below [@bookdown1; @bookdown2; @knitr1; @knitr2; @knitr3; @kableextra].
```{r eval=FALSE, include=TRUE}
install.packages(c("rmarkdown", "knitr", "kableExtra",
                   "stargazer", "modelsummary", "knitr", "gt"))
```
* ...download the 4 input files we created --- `paper.qmd`, `references.bib`, `data.csv` and `american-sociological-association.csl` --- from [this folder](https://drive.google.com/drive/folders/1rWzj-Bu1EKqkSuE1gaFzduBJhzlThpkw?usp=sharing). Ignore the other files.
* ...also download the 4 styling files we created: `wp_paged.html`, `wp.css`, `wp-fonts.css` and `wp-pages.css`.
* ...store all 8 files from above together in one folder (and use this folder as your working directory later on)
* ...learn R and read about the other underlying components namely [Markdown](https://en.wikipedia.org/wiki/Markdown), [R Markdown](https://rmarkdown.rstudio.com/lesson-1.html) and [Pagedown](https://pagedown.rbind.io/).
* ...pagedown comes with several Rmd-templates (presentations, poster, thesis, etc.) and via this review we provide another template for a working paper style. If however you want to modify single aspects or create your own template, you will need to at least gain some basic skills in [CSS](https://www.w3schools.com/css/) and [HTML](https://www.w3schools.com/html/).

# Basics: Input files, output files and the YAML header

All the files you need to produce the present PDF file are:

1. the input files:

* `paper.qmd` (the underlying R Markdown file).
* `references.bib` (the bibliography).
    + I use paperpile to manage my references and export the `.bib` file into the folder that contains my `.qmd` file.
* `data.csv` (some raw data).
* `american-sociological-association.csl` (defines the style of your bibliography).^[You can download various citation style files from this webpage: https://github.com/citation-style-language/styles.]

2. the "styling" files:

Basically, these are files you will need to specify in the YAML of your rmd-file, so that R and ultimately pagedown recognizes the certain style you want to achieve for your document. With using our templates, you will create a document that has the "look" of a working paper (we based our files on the [jss_paged pagedown format](https://github.com/rstudio/pagedown#journal-of-statistical-software-article-pagedownjss_paged)).

* `wp_paged.html` (based on `jss_paged.html`)
* `wp.css`
* `wp-fonts.css`
* `wp-pages.css`

Take `paper.qmd` (the underlying R Markdown file of this pdf) and have a look at the YAML (line \#18 - \#22) to see how to specifiy these files.
Basically, what happens here is that within the [jss_paged function](https://rdrr.io/cran/pagedown/man/jss_paged.html) we additionally specify that we want to use custom CSS and custom HTML.

[Download these files](https://drive.google.com/drive/folders/1rWzj-Bu1EKqkSuE1gaFzduBJhzlThpkw?usp=sharing) and save them into a folder. Close R/Rstudio and directly open `paper_pagedown.qmd` with RStudio. Doing so assures that the working directory is set to the folder that contains `paper.qmd` and the other files.^[You can always check your working directory in R with `getwd()`.]

Once you run/compile the `paper.qmd` file in Rstudio it creates a output file called `paper_pagedown.html`.

By using pagedown's `chrome_print` function in the YAML (line \#25) your html based web page will be printed to `paper_pagedown.pdf` (the one you are reading right now).

Both outputs will be saved in your working directory.

# Referencing within your document

To see how referencing works simply see the different examples for figures, tables and sections below. For instance in Section \@ref(sec:tables) you can find different ways of referencing tables. The code of the underlying `paper.qmd` will show you how I referenced Section \@ref(sec:tables) right here namely with '`Section \@ref(sec:tables)`'.

# Software versioning

Software changes and gets updated, especially with an active developer community like that of R. Luckily you can always access [old versions of R](https://cran.r-project.org/bin/windows/base/old/) and old version of R packages in [the archive](https://cran.r-project.org/src/contrib/Archive/). In the archive you need to choose a particular package, e.g dplyr and search for the right version, e.g., `dplyr_0.2.tar.gz`. Then insert the path in the following function: `install.packages("https://....../dplyr_0.2.tar.gz", repos=NULL, type="source")`. Ideally, however, results will be simply reproducible in the most current R and package versions.

I would recommend to use the command below and simply add it to the appendix as I did here in Appendix \@ref(sec:rsessioninfo). This will make sure you always inform the reader about the package versions your relied on in your paper. For more advanced tools see [packrat](https://rstudio.github.io/packrat/).

```{r fig-versioning, echo=TRUE, eval=FALSE}
cat(paste("#", capture.output(sessionInfo()), "\n", collapse ="")) 
  # or use message() instead of cat()
```

# Data
## Import

Generally, code is evaluated by inserting regular code chunks.

```{r}
x <- 1:10
x
```

Below we import an exemplary dataset (you can find `data.csv` in the folder with the other files).

```{r, echo=T, results="raw"}
data <- read.csv("data.csv")
head(data)
```

## Putting your entire data into the .qmd file
Applying the function `dput()` to an object gives you the code needed to reproduce that object. So you could paste that code into your `.qmd` file if you don't want to have extra data files. This makes sense were data files are small.

```{r}
dput(data[1:5,]) # here we only take a subset
```

You can then insert the dput output in an rchunk in your `.qmd` as below.

```{r eval=FALSE, include=TRUE, echo=TRUE}
data <- structure(list(Fertility = c(80.2, 83.1, 92.5, 85.8, 76.9), Agriculture = c(17, 
45.1, 39.7, 36.5, 43.5), Examination = c(15L, 6L, 5L, 12L, 17L
), Education = c(12L, 9L, 5L, 7L, 15L), Catholic = c(9.96, 84.84, 
93.4, 33.77, 5.16), Infant.Mortality = c(22.2, 22.2, 20.2, 20.3, 
20.6)), class = "data.frame", row.names = c(NA, -5L))
```

# Tables {#sec:tables}

Producing good tables and referencing these tables within a R Markdown document has been a hassle but got much better. Examples that you may use are shown below.

## Tables with kable() and kable_styling()

A great function is `kable()` (`knitr` package) in combination with `kableExtra`. @tbl-2 provides an example. To reference the table produced by the chunk you need to add `#| label: tbl-x` at the start of the chunk, i.e., `#| label: tbl-2` and would reference it by adding "`@tbl-2`" in your text.

 {{

```{r table-2, echo=TRUE}
#| label: tbl-2
#| tbl-cap: "Summary: Numeric variables"

library(knitr)
library(kableExtra)

kable(swiss[1:5,], row.names = TRUE, 
      caption = 'Table with kable() and kablestyling()', 
      format = "html", booktabs = T) %>%
        kable_styling(full_width = T, 
                      latex_options = c("striped", 
                                        "scale_down",
                                        "HOLD_position"),
                      font_size = 11)

```

## Tables with modelsummary

The `modelsummary` package provides a variety of tables and plots to summarize statistical models and data in R. Modellsummary plots and tables are highly customizable and they can be saved to almost all formats, e.g., HTML, PDF and Markdown. This makes ist especially easy to embed them in dynamic documents. Please look at the package's extensive [documentation](https://vincentarelbundock.github.io/modelsummary/index.html) where they also show examples for almost any plot or table you might be looking for. In this template we demonstrate an example for modelsummary's `datasummary` function. `Datasummary` creates frequency tables, crosstab tables, correlation tables, balance tables and many **more**.

### Summarize numeric variables

@tbl-3 shows a summary table for numeric variables.

```{r table-3, echo=TRUE}
#| label: tbl-3
#| tbl-cap: "Summary: Numeric variables"

library(modelsummary)
datasummary_skim(swiss, 
                 type="numeric", 
                 histogram=T, 
                 title = "Summary: Numeric variables")
```

### Summarize categorical variables

@tbl-4 shows a summary table for categorical variables.

```{r table-4, echo=TRUE}
#| label: tbl-4
#| tbl-cap: "Summary: Categorical variables"

# Create categorical variables
swiss$Education_cat <- cut(swiss$Education, 
                   breaks=c(-Inf, 6, 12, Inf), 
                   labels=c("low","middle","high"))
swiss$Infant.Mortality_cat <- cut(swiss$Infant.Mortality, 
                   breaks=c(-Inf, 18.15, 21.70, Inf), 
                   labels=c("low","middle","high"))

library(flextable)
tab_cat <- datasummary_skim(swiss, 
                            type="categorical", 
                            output = 'flextable')

# additionally we want to change the font, fontsize and spacing
library("gdtools")

library(dplyr)
tab_cat <- tab_cat %>%
  font(fontname="Times New Roman", part="header") %>%
  font(fontname="Times New Roman", j=1:4) %>% 
  fontsize(size=12, part="header") %>%
  fontsize(size=10, j=1:4) %>% 
  line_spacing(space = 0.3, part = "all")

tab_cat
```

## Regression table

@tbl-5 shows the output for a regression table. Make sure you name all the models you estimate (even if its 50) and explicitly refer to model names (M1, M2 etc.) in the text.

```{r}
#| label: tbl-5
#| tbl-cap: "Linear regression"

library(modelsummary)
M1 <- lm(Fertility ~ Education + Agriculture, data = swiss)
M2 <- lm(Fertility ~ Education + Catholic, data = swiss)
M3 <- lm(Fertility ~ Education + Infant.Mortality + Agriculture, data = swiss)
models <- list("M1" = M1, "M2" =  M2, "M3" = M3)

library(gt)
# additionally we want to change the font, font size and spacing
modelsummary(models,
             output = 'gt',
             notes = "Notes: some notes...") %>%
    tab_spanner(label = 'Dependent variable: Fertility', columns = 2:4) %>%
  tab_options(
    table.font.size = 10,
    data_row.padding = px(1),
    table.border.top.color = "white",
    heading.border.bottom.color = "black",
    row_group.border.top.color = "black",
    row_group.border.bottom.color = "white",
    table.border.bottom.color = "white",
    column_labels.border.top.color = "black",
    column_labels.border.bottom.color = "black",
    table_body.border.bottom.color = "black",
    table_body.hlines.color = "white"
  )

```

# Inline code & results

Reproduction reaches new heights when you work with inline code. For instance, you can automatize the display of certain coefficients within the text. An example is to include estimates, e.g., the coefficient of `dist` of the model we ran above. `r rinline("round(coef(M1)[2], 2)")` will insert the coefficient as follows: `r round(coef(M1)[2], 2)`. Or `r rinline("3 + 7")` will insert a `r 3 + 7` in the text.      
Inline code/results that depend on earlier objects in your document will automatically be updated once you change those objects. For instance, imagine a reviewer asks you to omit certain observations from your sample. You can simply do so in the beginning of your code and push play subsequently.. at time you might have to set `cache = FALSE` at the beginning so that all the code chunks are rerun.    
Researchers often avoid referring to results in-text etc. because you easily forget to change them when revising a manuscript. At the same it can make an article much more informative and easier to read, e.g., if you discuss a coefficient in the text you can directly show it in the section in which you discuss it. Inline code allows you to do just that. R Markdown allows you to that do so in a reproducible and automatized manner.

# Graphs

## R base graphs
Inserting figures can be slightly more complicated. Ideally, we would produce and insert them directly in the `.qmd` file. It's relatively simple to insert R base graphs as you can see in @fig1.

```{r}
#| label: fig1
#| fig-cap: "Scatterplot of Speed and Distance"

plot(swiss$Catholic, swiss$Fertility)
```

But it turns out that it doesn't always work so well.

## ggplot2 graphs
Same is true for ggplot2 as you can see in @fig2.

```{r}
#| label: fig2
#| fig-cap: "Miles per gallon according to the weight"

library(ggplot2)
ggplot(swiss, aes(x=Catholic, y=Fertility, shape=Education_cat)) + geom_point() +
  labs(x="Agriculture", y = "Fertility", 
       shape="Education") + theme_classic()
```

# Compiling the document

To view your paper, pagedown requires a web server (since it is based on paged.js)^[open-source library to paginate content in the browser]. By compiling a document, R Studio will display your HTML page through a local web server, i.e., paged.js will work in RStudio Viewer.

There are several options, depending on your intention:

- click on the `Knit` button in R Studio which by default will provide a HTML document in the RStudio viewer pane (the HTML will be stored in your working directory)
- use pagedown's `chrome_print` function in the YAML (uncomment line \#24 of this `Rmd` file) if additionally you want your HTML based web page to be printed to be PDF (the PDF will be stored in your working directory)
- to "live"-preview your pages do not click on the `Knit` button but use the **xaringan** [@xaringan] RStudio add-in _Infinite Moon Reader_. You can simply call the function `xaringan::inf_mr()` (within your console). This will launch a local web server via the **servr** package (Xie 2021a) and display your pages in the RStudio viewer. Each time you save your document (_Ctrl+S_) xaringan updates your pages in the viewer.

- If you use the option `self_contained: false` (see line \#21 of this `Rmd` file) (change to true for a self-contained document, but it'll be a litte slower for Pandoc to render), don't click on the `Knit` button in RStudio. Use instead the **xaringan** [@xaringan] RStudio add-in _Infinite Moon Reader_.

# Good practices for reproducibility
Every researcher has his own optimized setup. Currently we would recommend the following:

* Keep all files of your project (that matter for producing the PDF) in one folder without subfolders. You can zip and directly upload that folder to the [Harvard dataverse](https://dataverse.harvard.edu/).^[Another good folder setup would be to store all files needed as input files for the R Markdown manuscript in a subfolder called "input" and all output files that are produced apart from paper.html and paper.pdf in a subfolder called "output".]
* Make sure that filenames have a logic to them.
    + Main file with text/code: "paper.qmd", "report.qmd" 
    + Data files: "data_xxxxxx.*"
    + Image files: "fig_xxxxxx.*"
    + Tables files: "table_xxxx.*"
    + etc.
    + Ideally, your filenames will correspond to the names in the paper. For instance, Figure 1 in the paper may have a corresponding file called `fig_1_xxxxx.pdf`.
* Use the document outline in R studio (Ctrl + Shift + O) when you work with R Markdown.
* Name rchunks according to what they do or produce:
    + "`fig-...`" for chunks producing figures
    + "`table-...`" for chunks producing tables
    + "`model-...`" for chunks producing model estimates
    + "`import-...`" for chunks importing data
    + "`recoding-...`" for chunks in which data is recoded
* Use "really" informative variable names:
    + Q: What do you think does the variable *trstep* measure? It actually measures trust in the European parliament. 
        + How could we call this variable instead? Yes, `trust.european.parliament` which is longer but will probably be understood by another researcher in 50 years. 
    + If your setup is truly reproducible you will probably re-use the variable names that you generate as variable names in the tables you produce. Hence, there is an incentive to use good names.
* Use unique identifiers in the final R Markdown document paper.qmd that you upload:
    + Think of someone who wants to produce Figure 1/Model 1/Table 1 in your paper but doesn't find it in your code...
        + Name the chunks "fig-1", "fig-2" as the are named in the published paper.
        + Name the chunks that produce tables "table-1", "table-2" etc. as they are named in the published paper.
        + Name your statistical models in your R code "M1", "M2" as they are named in the published paper.

# Additional tricks for publishing
* Make your script anonymous
    + Simply put a `<!-- ... -->` around any identifying information, e.g., author names, title footnote etc.
* Counting words
    + Use adobe acrobat (commerical software) to convert your file to a word file. Then open in word and delete all the parts that shouldn't go into the word count. The word count is displayed in the lower right.
    + Use an one of the online services to count your words (search for "pdf word count")
* Appendix: You can change the numbering format for the appendix in the rmd file
    + What is still not possible in this document is to automatically have separate reference sections for paper and appendix.
* Journals may require you to use their tex style: Sometimes you can simply use their template in your rmarkdown file. See [here](https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/LDUMNY) for a PLOS one example.

# Citation styles
If your study needs to follow a particular citation style, you can set the corresponding style in the header of your `.qmd` document. To do so you have to download the corresponding `.csl` file.

In the present document we use the style of the American Sociological Association and set it in the preamble with `csl: american-sociological-association.csl`. However, you also need to download the respective `.csl` file from the following github page: https://github.com/citation-style-language/styles and copy it into your working directory for it to work.

The github directory contains a wide variety of citation style files depending on what discipline you work in.

\newpage

# References {-}

::: {#refs}

:::

<!--  to generate a citation entry for BibTeX, you can pass the returned object of citation() to toBibtex(), e.g. toBibtex(citation("xaringan")). Then  copy the output to a .bib file and add a unique citation key -->
<!-- https://bookdown.org/yihui/rmarkdown-cookbook/write-bib.html -->

\newpage

# Online appendix {-}

## Attach R session info in appendix {#sec:rsessioninfo}

Since R and R packages are constantly evolving you might want to add the R session info that contains information on the R version as well as the packages that are loaded.

```{r echo=FALSE}
print(sessionInfo(), local = FALSE)
```

## All the code in the paper

To simply attach all the code you used in the PDF file in the appendix see the R chunk in the underlying `.qmd` file:

```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE}
```

### 2023-02-17 paper.tex
First edits to past document.

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{241,243,245}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.40,0.45,0.13}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\BuiltInTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\ExtensionTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.28,0.35,0.67}{#1}}
\newcommand{\ImportTok}[1]{\textcolor[rgb]{0.00,0.46,0.62}{#1}}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\NormalTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.68,0.00,0.00}{#1}}
\newcommand{\RegionMarkerTok}[1]{\textcolor[rgb]{0.00,0.23,0.31}{#1}}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.07,0.07,0.07}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.13,0.47,0.30}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.37,0.37,0.37}{\textit{#1}}}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{siunitx}

  \newcolumntype{d}{S[
    input-open-uncertainty=,
    input-close-uncertainty=,
    parse-numbers = false,
    table-align-text-pre=false,
    table-align-text-post=false
  ]}

\usepackage{fontspec}
\usepackage{multicol}
\usepackage{hhline}
\newlength\Oldarrayrulewidth
\newlength\Oldtabcolsep
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{caption}
\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Writing a reproducible paper with Quarto},
  pdfauthor={Paul C. Bauer, Camille Landesvatter},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Writing a reproducible paper with Quarto}
\author{Paul C. Bauer, Camille Landesvatter}
\date{5/4/18}

\begin{document}
\maketitle
\begin{abstract}
The present paper provides a template to write a reproducible scientific
paper with R Markdown and Pagedown.\footnote{Based on an earlier R
  Markdown template that uses Latex and can be downloaded under
  https://osf.io/q395s (see Bauer 2018).} Below we outline some of the
``tricks''/code (e.g., referencing tables, sections etc.) we had to
figure out to produce this document. The underlying files which produce
this document can be
\href{https://drive.google.com/drive/folders/1rWzj-Bu1EKqkSuE1gaFzduBJhzlThpkw?usp=sharing}{downloaded
here}. Importantly, we also provide different CSS and HTML files that
can be used to achieve a pdf output with the look of a ``working
paper''. We are convinced that in the future there will be many
improvements and developments with regards to RStudio, R markdown and
Pagedown. We intend to update this file when we discover more convenient
code. You can follow any updates on the
\href{https://github.com/paulcbauer/Writing_a_reproducable_paper_with_quarto/}{github
repository}.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, enhanced, interior hidden, borderline west={3pt}{0pt}{shadecolor}, frame hidden, sharp corners, breakable]}{\end{tcolorbox}}\fi

\renewcommand*\contentsname{Table of contents}
{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{3}
\tableofcontents
}
\hypertarget{why-reproducible-research-in-r}{%
\section{Why reproducible research (in
R)?}\label{why-reproducible-research-in-r}}

Some arguments\ldots{}

\begin{itemize}
\item
  \textbf{Access}: Research is normally funded by taxpayers (researchers
  are also taxpayers). Hence, it should be freely accessible to everyone
  without any barriers, e.g., without requiring commercial software.
  Importantly, researchers from developing countries are even more
  dependent on free access to knowledge (Kirsop and Chan 2005).
\item
  \textbf{Reproducability}: Even if you have written a study and
  analyzed the data yourself you will forget what you did after a few
  months. A fully reproducible setup will help you to trace back your
  own steps. Obviously, the same is true for other researchers who may
  want to understand your work and built on it. It may sound like a joke
  but why not aim for a document that can be used to reproduce your
  findings in 500 years.
\item
  \textbf{Errors}: Manual steps in data analysis (e.g., manually
  copy/pasting values into a table etc.) may introduce errors. R
  Markdown allows you to \textbf{automatize} such steps and/or avoid
  them.
\item
  \textbf{Revisions}: Revising a paper takes much less time if you have
  all the code you need in one place, i.e., one \texttt{.rmd} file. For
  instance, if you decide to exclude a subset of your data you simply
  need to insert one line of your code at the beginning and everything
  is rebuilt/re-estimated automatically.
\end{itemize}

\hypertarget{why-quarto}{%
\section{Why Quarto?}\label{why-quarto}}

Formatting text as PDF is probably one of the most widespread standards
in the scientific community, especially when it comes to submitting
papers and similar documents. The traditional way to well-formatted and
good-looking PDFs is often through LaTeX or Word. However, if you have
spent hours and hours debugging latex code (or getting it to run) you
may be on the lookout for something new.

The fairly new \texttt{pagedown} R package takes a completely new
approach. While the main purpose of pagedown is to create high-quality
PDFs, the idea is to take advantage of modern web technologies
(HTML/JSS/Javascript) with which one can design web pages and eventually
print those to PDF.

While web pages are usually single-page scrollable documents, pagedown
uses the JavaScript library \texttt{Paged.js} which allows documents to
be paginated with elements like headers, footers and everything a
readable scientific paper will need. Additionally, pagedown documents
are based on R Markdown. In our view, Pagedown and the underlying
technology may replace Latex in the long run. In the near future it
should also be possible to produce a PDF with static graph and an
equivalent html with interactive graps (see dicsussion
\href{https://github.com/rstudio/pagedown/pull/87}{here}).

\hypertarget{prerequesites}{%
\section{Prerequesites}\label{prerequesites}}

We assume that you are using R on a day-to-day basis and you may have
even started to work in R Markdown. If you don't know what R Markdown is
there are many great resources that you should use (e.g.~watch this
\href{https://vimeo.com/178485416}{short video}). An older template
{[}see Bauer (2018); https://osf.io/q395s{]} on which this newer
template is based, may provide a quick entry point to writing a
reproducible with R Markdown and Latex.

Based on R Markdown, Pagedown allows you to create custom and
well-formatted (paged) HTML Documents. For a comprehensive overview
watch
\href{https://www.rstudio.com/resources/rstudioconf-2019/pagedown-creating-beautiful-pdfs-with-r-markdown-and-css/}{this
video} which is a record of a talk introducing \texttt{pagedown} given
by Yihui Xie (who in addition to Romain Lesur developed the
\texttt{pagedown} package). If you are not in a video watching mood find
the slides
\href{https://slides.yihui.org/2019-rstudio-conf-pagedown.html\#1}{here}.

Then\ldots{}

\begin{itemize}
\tightlist
\item
  \ldots install \href{https://www.r-project.org/}{R} and
  \href{https://www.rstudio.com/}{Rstudio} (most recent versions) (R
  Core Team 2017; RStudio Team 2015).
\item
  \ldots install the ``pagedown''-package from github using the code
  below (Xie and Lesur 2021; Xie et al. 2021).
\end{itemize}

\begin{itemize}
\tightlist
\item
  \ldots also install the packages below using the code below (Xie 2017,
  2016, 2018, 2015, 2014; Zhu 2017).
\end{itemize}

\begin{itemize}
\tightlist
\item
  \ldots download the 4 input files we created --- \texttt{paper.rmd},
  \texttt{references.bib}, \texttt{data.csv} and
  \texttt{american-sociological-association.csl} --- from
  \href{https://drive.google.com/drive/folders/1rWzj-Bu1EKqkSuE1gaFzduBJhzlThpkw?usp=sharing}{this
  folder}. Ignore the other files.
\item
  \ldots also download the 4 styling files we created:
  \texttt{wp\_paged.html}, \texttt{wp.css}, \texttt{wp-fonts.css} and
  \texttt{wp-pages.css}.
\item
  \ldots store all 8 files from above together in one folder (and use
  this folder as your working directory later on)
\item
  \ldots learn R and read about the other underlying components namely
  \href{https://en.wikipedia.org/wiki/Markdown}{Markdown},
  \href{https://rmarkdown.rstudio.com/lesson-1.html}{R Markdown} and
  \href{https://pagedown.rbind.io/}{Pagedown}.
\item
  \ldots pagedown comes with several Rmd-templates (presentations,
  poster, thesis, etc.) and via this review we provide another template
  for a working paper style. If however you want to modify single
  aspects or create your own template, you will need to at least gain
  some basic skills in \href{https://www.w3schools.com/css/}{CSS} and
  \href{https://www.w3schools.com/html/}{HTML}.
\end{itemize}

\hypertarget{basics-input-files-output-files-and-the-yaml-header}{%
\section{Basics: Input files, output files and the YAML
header}\label{basics-input-files-output-files-and-the-yaml-header}}

All the files you need to produce the present PDF file are:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  the input files:
\end{enumerate}

\begin{itemize}
\tightlist
\item
  \texttt{paper.rmd} (the underlying R Markdown file).
\item
  \texttt{references.bib} (the bibliography).

  \begin{itemize}
  \tightlist
  \item
    I use paperpile to manage my references and export the \texttt{.bib}
    file into the folder that contains my \texttt{.rmd} file.
  \end{itemize}
\item
  \texttt{data.csv} (some raw data).
\item
  \texttt{american-sociological-association.csl} (defines the style of
  your bibliography).\footnote{You can download various citation style
    files from this webpage:
    https://github.com/citation-style-language/styles.}
\end{itemize}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{1}
\tightlist
\item
  the ``styling'' files:
\end{enumerate}

Basically, these are files you will need to specify in the YAML of your
rmd-file, so that R and ultimately pagedown recognizes the certain style
you want to achieve for your document. With using our templates, you
will create a document that has the ``look'' of a working paper (we
based our files on the
\href{https://github.com/rstudio/pagedown\#journal-of-statistical-software-article-pagedownjss_paged}{jss\_paged
pagedown format}).

\begin{itemize}
\tightlist
\item
  \texttt{wp\_paged.html} (based on \texttt{jss\_paged.html})
\item
  \texttt{wp.css}
\item
  \texttt{wp-fonts.css}
\item
  \texttt{wp-pages.css}
\end{itemize}

Take \texttt{paper.rmd} (the underlying R Markdown file of this pdf) and
have a look at the YAML (line \#18 - \#22) to see how to specifiy these
files. Basically, what happens here is that within the
\href{https://rdrr.io/cran/pagedown/man/jss_paged.html}{jss\_paged
function} we additionally specify that we want to use custom CSS and
custom HTML.

\href{https://drive.google.com/drive/folders/1rWzj-Bu1EKqkSuE1gaFzduBJhzlThpkw?usp=sharing}{Download
these files} and save them into a folder. Close R/Rstudio and directly
open \texttt{paper\_pagedown.rmd} with RStudio. Doing so assures that
the working directory is set to the folder that contains
\texttt{paper.rmd} and the other files.\footnote{You can always check
  your working directory in R with \texttt{getwd()}.}

Once you run/compile the \texttt{paper.rmd} file in Rstudio it creates a
output file called \texttt{paper\_pagedown.html}.

By using pagedown's \texttt{chrome\_print} function in the YAML (line
\#25) your html based web page will be printed to
\texttt{paper\_pagedown.pdf} (the one you are reading right now).

Both outputs will be saved in your working directory.

\hypertarget{referencing-within-your-document}{%
\section{Referencing within your
document}\label{referencing-within-your-document}}

To see how referencing works simply see the different examples for
figures, tables and sections below. For instance in Section
@ref(sec:tables) you can find different ways of referencing tables. The
code of the underlying \texttt{paper.rmd} will show you how I referenced
Section @ref(sec:tables) right here namely with
`\texttt{Section\ \textbackslash{}@ref(sec:tables)}'.

\hypertarget{software-versioning}{%
\section{Software versioning}\label{software-versioning}}

Software changes and gets updated, especially with an active developer
community like that of R. Luckily you can always access
\href{https://cran.r-project.org/bin/windows/base/old/}{old versions of
R} and old version of R packages in
\href{https://cran.r-project.org/src/contrib/Archive/}{the archive}. In
the archive you need to choose a particular package, e.g dplyr and
search for the right version, e.g., \texttt{dplyr\_0.2.tar.gz}. Then
insert the path in the following function:
\texttt{install.packages("https://....../dplyr\_0.2.tar.gz",\ repos=NULL,\ type="source")}.
Ideally, however, results will be simply reproducible in the most
current R and package versions.

I would recommend to use the command below and simply add it to the
appendix as I did here in Appendix @ref(sec:rsessioninfo). This will
make sure you always inform the reader about the package versions your
relied on in your paper. For more advanced tools see
\href{https://rstudio.github.io/packrat/}{packrat}.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{cat}\NormalTok{(}\FunctionTok{paste}\NormalTok{(}\StringTok{"\#"}\NormalTok{, }\FunctionTok{capture.output}\NormalTok{(}\FunctionTok{sessionInfo}\NormalTok{()), }\StringTok{"}\SpecialCharTok{\textbackslash{}n}\StringTok{"}\NormalTok{, }\AttributeTok{collapse =}\StringTok{""}\NormalTok{)) }
  \CommentTok{\# or use message() instead of cat()}
\end{Highlighting}
\end{Shaded}

\hypertarget{data}{%
\section{Data}\label{data}}

\hypertarget{import}{%
\subsection{Import}\label{import}}

Generally, code is evaluated by inserting regular \texttt{R\ Markdown}
blocks.

\begin{verbatim}
 [1]  1  2  3  4  5  6  7  8  9 10
\end{verbatim}

Below we import an exemplary dataset (you can find \texttt{data.csv} in
the folder with the other files).

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{data }\OtherTok{\textless{}{-}} \FunctionTok{read.csv}\NormalTok{(}\StringTok{"data.csv"}\NormalTok{)}
\FunctionTok{head}\NormalTok{(data)}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
  Fertility Agriculture Examination Education Catholic Infant.Mortality
1      80.2        17.0          15        12     9.96             22.2
2      83.1        45.1           6         9    84.84             22.2
3      92.5        39.7           5         5    93.40             20.2
4      85.8        36.5          12         7    33.77             20.3
5      76.9        43.5          17        15     5.16             20.6
6      76.1        35.3           9         7    90.57             26.6
\end{verbatim}

\hypertarget{putting-your-entire-data-into-the-.rmd-file}{%
\subsection{Putting your entire data into the .rmd
file}\label{putting-your-entire-data-into-the-.rmd-file}}

Applying the function \texttt{dput()} to an object gives you the code
needed to reproduce that object. So you could paste that code into your
\texttt{.rmd} file if you don't want to have extra data files. This
makes sense were data files are small.

\begin{verbatim}
structure(list(Fertility = c(80.2, 83.1, 92.5, 85.8, 76.9), Agriculture = c(17, 
45.1, 39.7, 36.5, 43.5), Examination = c(15L, 6L, 5L, 12L, 17L
), Education = c(12L, 9L, 5L, 7L, 15L), Catholic = c(9.96, 84.84, 
93.4, 33.77, 5.16), Infant.Mortality = c(22.2, 22.2, 20.2, 20.3, 
20.6)), row.names = c(NA, 5L), class = "data.frame")
\end{verbatim}

You can then insert the dput output in your \texttt{.rmd} as below.

\hypertarget{sec:tables}{%
\section{Tables}\label{sec:tables}}

Producing good tables and referencing these tables within a R Markdown
document has been a hassle but got much better. Examples that you may
use are shown below.

\hypertarget{tables-with-kable-and-kable_styling}{%
\subsection{Tables with kable() and
kable\_styling()}\label{tables-with-kable-and-kable_styling}}

A great function is \texttt{kable()} (\texttt{knitr} package) in
combination with \texttt{kableExtra}. Table @ref(tab:table-2) provides
an example. To reference the table produced by the chunk you need to add
´tab:´ to the chunk name, i.e., ´tab:table-2´ and would reference it by
adding ``\texttt{Table\ \textbackslash{}@ref(tab:table-2)}'' in your
text.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(knitr)}
\FunctionTok{library}\NormalTok{(kableExtra)}

\FunctionTok{kable}\NormalTok{(swiss[}\DecValTok{1}\SpecialCharTok{:}\DecValTok{10}\NormalTok{,], }\AttributeTok{row.names =} \ConstantTok{TRUE}\NormalTok{, }
      \AttributeTok{caption =} \StringTok{\textquotesingle{}Table with kable() and kablestyling()\textquotesingle{}}\NormalTok{, }
      \AttributeTok{format =} \StringTok{"html"}\NormalTok{, }\AttributeTok{booktabs =}\NormalTok{ T) }\SpecialCharTok{\%\textgreater{}\%}
        \FunctionTok{kable\_styling}\NormalTok{(}\AttributeTok{full\_width =}\NormalTok{ T, }
                      \AttributeTok{latex\_options =} \FunctionTok{c}\NormalTok{(}\StringTok{"striped"}\NormalTok{, }
                                        \StringTok{"scale\_down"}\NormalTok{,}
                                        \StringTok{"HOLD\_position"}\NormalTok{),}
                      \AttributeTok{font\_size =} \DecValTok{10}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{longtable}[]{@{}lrrrrrr@{}}
\caption{Table with kable() and kablestyling()}\tabularnewline
\toprule()
& Fertility & Agriculture & Examination & Education & Catholic &
Infant.Mortality \\
\midrule()
\endfirsthead
\toprule()
& Fertility & Agriculture & Examination & Education & Catholic &
Infant.Mortality \\
\midrule()
\endhead
Courtelary & 80.2 & 17.0 & 15 & 12 & 9.96 & 22.2 \\
Delemont & 83.1 & 45.1 & 6 & 9 & 84.84 & 22.2 \\
Franches-Mnt & 92.5 & 39.7 & 5 & 5 & 93.40 & 20.2 \\
Moutier & 85.8 & 36.5 & 12 & 7 & 33.77 & 20.3 \\
Neuveville & 76.9 & 43.5 & 17 & 15 & 5.16 & 20.6 \\
Porrentruy & 76.1 & 35.3 & 9 & 7 & 90.57 & 26.6 \\
Broye & 83.8 & 70.2 & 16 & 7 & 92.85 & 23.6 \\
Glane & 92.4 & 67.8 & 14 & 8 & 97.16 & 24.9 \\
Gruyere & 82.4 & 53.3 & 12 & 7 & 97.67 & 21.0 \\
Sarine & 82.9 & 45.2 & 16 & 13 & 91.38 & 24.4 \\
\bottomrule()
\end{longtable}

\hypertarget{tables-with-modelsummary}{%
\subsection{Tables with modelsummary}\label{tables-with-modelsummary}}

The \texttt{modelsummary} package provides a variety of tables and plots
to summarize statistical models and data in R. Modellsummary plots and
tables are highly customizable and they can be saved to almost all
formats, e.g., HTML, PDF and Markdown. This makes ist especially easy to
embed them in dynamic documents. Please look at the package's extensive
\href{https://vincentarelbundock.github.io/modelsummary/index.html}{documentation}
where they also show examples for almost any plot or table you might be
looking for. In this template we demonstrate an example for
modelsummary's \texttt{datasummary} function. \texttt{Datasummary}
creates frequency tables, crosstab tables, correlation tables, balance
tables and many \textbf{more}.

\hypertarget{summarize-numeric-variables}{%
\subsubsection{Summarize numeric
variables}\label{summarize-numeric-variables}}

Table @ref(tab:table-3) shows a summary table for numeric variables.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(modelsummary)}
\FunctionTok{datasummary\_skim}\NormalTok{(swiss, }
                 \AttributeTok{type=}\StringTok{"numeric"}\NormalTok{, }
                 \AttributeTok{histogram=}\NormalTok{T, }
                 \AttributeTok{title =} \StringTok{"Summary: Numeric variables"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{table}

\caption{Summary: Numeric variables}
\centering
\begin{tabular}[t]{lrrrrrrr>{}r}
\toprule
  & Unique (\#) & Missing (\%) & Mean & SD & Min & Median & Max &   \\
\midrule
Fertility & 46 & 0 & \num{70.1} & \num{12.5} & \num{35.0} & \num{70.4} & \num{92.5} & \includegraphics[width=0.67in, height=0.17in]{C:/Users/Paul/Google Drive/1-Research/2023_Writing_a_reproducable_paper_with_quarto/paper_files/figure-latex/hist_326c5c73d53.pdf}\\
Agriculture & 47 & 0 & \num{50.7} & \num{22.7} & \num{1.2} & \num{54.1} & \num{89.7} & \includegraphics[width=0.67in, height=0.17in]{C:/Users/Paul/Google Drive/1-Research/2023_Writing_a_reproducable_paper_with_quarto/paper_files/figure-latex/hist_326cfb12fe8.pdf}\\
Examination & 22 & 0 & \num{16.5} & \num{8.0} & \num{3.0} & \num{16.0} & \num{37.0} & \includegraphics[width=0.67in, height=0.17in]{C:/Users/Paul/Google Drive/1-Research/2023_Writing_a_reproducable_paper_with_quarto/paper_files/figure-latex/hist_326c1fe56795.pdf}\\
Education & 19 & 0 & \num{11.0} & \num{9.6} & \num{1.0} & \num{8.0} & \num{53.0} & \includegraphics[width=0.67in, height=0.17in]{C:/Users/Paul/Google Drive/1-Research/2023_Writing_a_reproducable_paper_with_quarto/paper_files/figure-latex/hist_326c7ee65a18.pdf}\\
Catholic & 46 & 0 & \num{41.1} & \num{41.7} & \num{2.1} & \num{15.1} & \num{100.0} & \includegraphics[width=0.67in, height=0.17in]{C:/Users/Paul/Google Drive/1-Research/2023_Writing_a_reproducable_paper_with_quarto/paper_files/figure-latex/hist_326c121e2da6.pdf}\\
Infant.Mortality & 37 & 0 & \num{19.9} & \num{2.9} & \num{10.8} & \num{20.0} & \num{26.6} & \includegraphics[width=0.67in, height=0.17in]{C:/Users/Paul/Google Drive/1-Research/2023_Writing_a_reproducable_paper_with_quarto/paper_files/figure-latex/hist_326c64e452d5.pdf}\\
\bottomrule
\end{tabular}
\end{table}

\hypertarget{summarize-categorical-variables}{%
\subsubsection{Summarize categorical
variables}\label{summarize-categorical-variables}}

Table @ref(tab:table-4) shows a summary table for categorical variables.

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# Create categorical variables}
\NormalTok{swiss}\SpecialCharTok{$}\NormalTok{Education\_cat }\OtherTok{\textless{}{-}} \FunctionTok{cut}\NormalTok{(swiss}\SpecialCharTok{$}\NormalTok{Education, }
                   \AttributeTok{breaks=}\FunctionTok{c}\NormalTok{(}\SpecialCharTok{{-}}\ConstantTok{Inf}\NormalTok{, }\DecValTok{6}\NormalTok{, }\DecValTok{12}\NormalTok{, }\ConstantTok{Inf}\NormalTok{), }
                   \AttributeTok{labels=}\FunctionTok{c}\NormalTok{(}\StringTok{"low"}\NormalTok{,}\StringTok{"middle"}\NormalTok{,}\StringTok{"high"}\NormalTok{))}
\NormalTok{swiss}\SpecialCharTok{$}\NormalTok{Infant.Mortality\_cat }\OtherTok{\textless{}{-}} \FunctionTok{cut}\NormalTok{(swiss}\SpecialCharTok{$}\NormalTok{Infant.Mortality, }
                   \AttributeTok{breaks=}\FunctionTok{c}\NormalTok{(}\SpecialCharTok{{-}}\ConstantTok{Inf}\NormalTok{, }\FloatTok{18.15}\NormalTok{, }\FloatTok{21.70}\NormalTok{, }\ConstantTok{Inf}\NormalTok{), }
                   \AttributeTok{labels=}\FunctionTok{c}\NormalTok{(}\StringTok{"low"}\NormalTok{,}\StringTok{"middle"}\NormalTok{,}\StringTok{"high"}\NormalTok{))}

\FunctionTok{library}\NormalTok{(flextable)}
\NormalTok{tab\_cat }\OtherTok{\textless{}{-}} \FunctionTok{datasummary\_skim}\NormalTok{(swiss, }
                            \AttributeTok{type=}\StringTok{"categorical"}\NormalTok{, }
                            \AttributeTok{title =} \StringTok{"Summary: Categorical variables"}\NormalTok{,}
                            \AttributeTok{output =} \StringTok{\textquotesingle{}flextable\textquotesingle{}}\NormalTok{)}

\CommentTok{\# additionally we want to change the font, fontsize and spacing}
\FunctionTok{library}\NormalTok{(}\StringTok{"gdtools"}\NormalTok{)}

\FunctionTok{library}\NormalTok{(dplyr)}
\NormalTok{tab\_cat }\OtherTok{\textless{}{-}}\NormalTok{ tab\_cat }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{font}\NormalTok{(}\AttributeTok{fontname=}\StringTok{"Times New Roman"}\NormalTok{, }\AttributeTok{part=}\StringTok{"header"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{font}\NormalTok{(}\AttributeTok{fontname=}\StringTok{"Times New Roman"}\NormalTok{, }\AttributeTok{j=}\DecValTok{1}\SpecialCharTok{:}\DecValTok{4}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{fontsize}\NormalTok{(}\AttributeTok{size=}\DecValTok{12}\NormalTok{, }\AttributeTok{part=}\StringTok{"header"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{fontsize}\NormalTok{(}\AttributeTok{size=}\DecValTok{10}\NormalTok{, }\AttributeTok{j=}\DecValTok{1}\SpecialCharTok{:}\DecValTok{4}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{line\_spacing}\NormalTok{(}\AttributeTok{space =} \FloatTok{0.3}\NormalTok{, }\AttributeTok{part =} \StringTok{"all"}\NormalTok{)}

\NormalTok{tab\_cat}
\end{Highlighting}
\end{Shaded}

\global\setlength{\Oldarrayrulewidth}{\arrayrulewidth}

\global\setlength{\Oldtabcolsep}{\tabcolsep}

\setlength{\tabcolsep}{0pt}

\renewcommand*{\arraystretch}{1.5}

\providecommand{\ascline}[3]{\noalign{\global\arrayrulewidth #1}\arrayrulecolor[HTML]{#2}\cline{#3}}

\begin{longtable}[c]{|p{0.75in}|p{0.75in}|p{0.75in}|p{0.75in}}

\caption{Summary: Categorical variables} \\

\ascline{2pt}{666666}{1-4}

\multicolumn{1}{>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{12}{4}\selectfont{\ }}} & \multicolumn{1}{>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{12}{4}\selectfont{\ \ }}} & \multicolumn{1}{>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{12}{4}\selectfont{N}}} & \multicolumn{1}{>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{12}{4}\selectfont{\%}}} \\

\ascline{2pt}{666666}{1-4}\endhead

\multicolumn{1}{>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{3}\selectfont{Education\_cat}}} & \multicolumn{1}{>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{3}\selectfont{low}}} & \multicolumn{1}{>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{3}\selectfont{14}}} & \multicolumn{1}{>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{3}\selectfont{29.8}}} \\

\multicolumn{1}{>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{3}\selectfont{}}} & \multicolumn{1}{>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{3}\selectfont{middle}}} & \multicolumn{1}{>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{3}\selectfont{22}}} & \multicolumn{1}{>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{3}\selectfont{46.8}}} \\

\multicolumn{1}{>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{3}\selectfont{}}} & \multicolumn{1}{>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{3}\selectfont{high}}} & \multicolumn{1}{>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{3}\selectfont{11}}} & \multicolumn{1}{>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{3}\selectfont{23.4}}} \\

\multicolumn{1}{>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{3}\selectfont{Infant.Mortality\_cat}}} & \multicolumn{1}{>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{3}\selectfont{low}}} & \multicolumn{1}{>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{3}\selectfont{12}}} & \multicolumn{1}{>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{3}\selectfont{25.5}}} \\

\multicolumn{1}{>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{3}\selectfont{}}} & \multicolumn{1}{>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{3}\selectfont{middle}}} & \multicolumn{1}{>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{3}\selectfont{23}}} & \multicolumn{1}{>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{3}\selectfont{48.9}}} \\

\multicolumn{1}{>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{3}\selectfont{}}} & \multicolumn{1}{>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{3}\selectfont{high}}} & \multicolumn{1}{>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{3}\selectfont{12}}} & \multicolumn{1}{>{\raggedright}p{\dimexpr 0.75in+0\tabcolsep}}{\textcolor[HTML]{000000}{\fontsize{10}{3}\selectfont{25.5}}} \\

\ascline{2pt}{666666}{1-4}

\end{longtable}

\arrayrulecolor[HTML]{000000}

\global\setlength{\arrayrulewidth}{\Oldarrayrulewidth}

\global\setlength{\tabcolsep}{\Oldtabcolsep}

\renewcommand*{\arraystretch}{1}

\hypertarget{regression-table}{%
\subsection{Regression table}\label{regression-table}}

Table @ref(tab:table-5) shows the output for a regression table. Make
sure you name all the models you estimate (even if its 50) and
explicitly refer to model names (M1, M2 etc.) in the text.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(modelsummary)}
\NormalTok{M1 }\OtherTok{\textless{}{-}} \FunctionTok{lm}\NormalTok{(Fertility }\SpecialCharTok{\textasciitilde{}}\NormalTok{ Education }\SpecialCharTok{+}\NormalTok{ Agriculture, }\AttributeTok{data =}\NormalTok{ swiss)}
\NormalTok{M2 }\OtherTok{\textless{}{-}} \FunctionTok{lm}\NormalTok{(Fertility }\SpecialCharTok{\textasciitilde{}}\NormalTok{ Education }\SpecialCharTok{+}\NormalTok{ Catholic, }\AttributeTok{data =}\NormalTok{ swiss)}
\NormalTok{M3 }\OtherTok{\textless{}{-}} \FunctionTok{lm}\NormalTok{(Fertility }\SpecialCharTok{\textasciitilde{}}\NormalTok{ Education }\SpecialCharTok{+}\NormalTok{ Infant.Mortality }\SpecialCharTok{+}\NormalTok{ Agriculture, }\AttributeTok{data =}\NormalTok{ swiss)}
\NormalTok{models }\OtherTok{\textless{}{-}} \FunctionTok{list}\NormalTok{(}\StringTok{"M1"} \OtherTok{=}\NormalTok{ M1, }\StringTok{"M2"} \OtherTok{=}\NormalTok{  M2, }\StringTok{"M3"} \OtherTok{=}\NormalTok{ M3)}

\FunctionTok{library}\NormalTok{(gt)}
\CommentTok{\# additionally we want to change the font, font size and spacing}
\FunctionTok{modelsummary}\NormalTok{(models,}
             \AttributeTok{title =} \StringTok{\textquotesingle{}Linear regression\textquotesingle{}}\NormalTok{,}
             \AttributeTok{output =} \StringTok{\textquotesingle{}gt\textquotesingle{}}\NormalTok{,}
             \AttributeTok{notes =} \StringTok{"Notes: some notes..."}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{tab\_spanner}\NormalTok{(}\AttributeTok{label =} \StringTok{\textquotesingle{}Dependent variable: Fertility\textquotesingle{}}\NormalTok{, }\AttributeTok{columns =} \DecValTok{2}\SpecialCharTok{:}\DecValTok{4}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{tab\_options}\NormalTok{(}
    \AttributeTok{table.font.size =} \DecValTok{10}\NormalTok{,}
    \AttributeTok{data\_row.padding =} \FunctionTok{px}\NormalTok{(}\DecValTok{1}\NormalTok{),}
    \AttributeTok{table.border.top.color =} \StringTok{"white"}\NormalTok{,}
    \AttributeTok{heading.border.bottom.color =} \StringTok{"black"}\NormalTok{,}
    \AttributeTok{row\_group.border.top.color =} \StringTok{"black"}\NormalTok{,}
    \AttributeTok{row\_group.border.bottom.color =} \StringTok{"white"}\NormalTok{,}
    \AttributeTok{table.border.bottom.color =} \StringTok{"white"}\NormalTok{,}
    \AttributeTok{column\_labels.border.top.color =} \StringTok{"black"}\NormalTok{,}
    \AttributeTok{column\_labels.border.bottom.color =} \StringTok{"black"}\NormalTok{,}
    \AttributeTok{table\_body.border.bottom.color =} \StringTok{"black"}\NormalTok{,}
    \AttributeTok{table\_body.hlines.color =} \StringTok{"white"}
\NormalTok{  )}
\end{Highlighting}
\end{Shaded}

\setlength{\LTpost}{0mm}
\begin{longtable}{lccc}
\toprule
 & \multicolumn{3}{c}{Dependent variable: Fertility} \\ 
\cmidrule(lr){2-4}
  & M1 & M2 & M3 \\ 
\midrule
(Intercept) & 84.080 & 74.234 & 51.101 \\ 
 & (5.782) & (2.352) & (10.995) \\ 
Education & -0.963 & -0.788 & -0.857 \\ 
 & (0.189) & (0.129) & (0.173) \\ 
Agriculture & -0.066 &  & -0.026 \\ 
 & (0.080) &  & (0.073) \\ 
Catholic &  & 0.111 &  \\ 
 &  & (0.030) &  \\ 
Infant.Mortality &  &  & 1.493 \\ 
 &  &  & (0.439) \\ 
Num.Obs. & 47 & 47 & 47 \\ 
R2 & 0.449 & 0.575 & 0.566 \\ 
R2 Adj. & 0.424 & 0.555 & 0.536 \\ 
AIC & 349.7 & 337.6 & 340.5 \\ 
BIC & 357.1 & 345.0 & 349.7 \\ 
Log.Lik. & -170.846 & -164.782 & -165.243 \\ 
RMSE & 9.17 & 8.06 & 8.14 \\ 
\bottomrule
\end{longtable}
\begin{minipage}{\linewidth}
Notes: some notes...\\
\end{minipage}

\hypertarget{inline-code-results}{%
\section{Inline code \& results}\label{inline-code-results}}

Reproduction reaches new heights when you work with inline code. For
instance, you can automatize the display of certain coefficients within
the text. An example is to include estimates, e.g., the coefficient of
\texttt{dist} of the model we ran above.
\texttt{\textasciigrave{}r\ round(coef(M1){[}2{]},\ 2)\textasciigrave{}}
will insert the coefficient as follows: -0.96. Or
\texttt{\textasciigrave{}r\ 3\ +\ 7\textasciigrave{}} will insert a 10
in the text.\\
Inline code/results that depend on earlier objects in your document will
automatically be updated once you change those objects. For instance,
imagine a reviewer asks you to omit certain observations from your
sample. You can simply do so in the beginning of your code and push play
subsequently.. at time you might have to set \texttt{cache\ =\ FALSE} at
the beginning so that all the code chunks are rerun.\\
Researchers often avoid referring to results in-text etc. because you
easily forget to change them when revising a manuscript. At the same it
can make an article much more informative and easier to read, e.g., if
you discuss a coefficient in the text you can directly show it in the
section in which you discuss it. Inline code allows you to do just that.
R Markdown allows you to that do so in a reproducible and automatized
manner.

\hypertarget{graphs}{%
\section{Graphs}\label{graphs}}

\hypertarget{r-base-graphs}{%
\subsection{R base graphs}\label{r-base-graphs}}

Inserting figures can be slightly more complicated. Ideally, we would
produce and insert them directly in the \texttt{.rmd} file. It's
relatively simple to insert R base graphs as you can see in Figure
@ref(fig:fig-1).

\begin{figure}[H]

{\centering \includegraphics{paper_files/figure-pdf/fig-1-1.pdf}

}

\caption{\label{fig-1}Scatterplot of Speed and Distance}

\end{figure}

But it turns out that it doesn't always work so well.

\hypertarget{ggplot2-graphs}{%
\subsection{ggplot2 graphs}\label{ggplot2-graphs}}

Same is true for ggplot2 as you can see in Figure @ref(fig:fig-2).

\begin{figure}[H]

{\centering \includegraphics{paper_files/figure-pdf/fig-2-1.pdf}

}

\caption{\label{fig-2}Miles per gallon according to the weight}

\end{figure}

\hypertarget{compiling-the-document}{%
\section{Compiling the document}\label{compiling-the-document}}

To view your paper, pagedown requires a web server (since it is based on
paged.js)\footnote{open-source library to paginate content in the
  browser}. By compiling a document, R Studio will display your HTML
page through a local web server, i.e., paged.js will work in RStudio
Viewer.

There are several options, depending on your intention:

\begin{itemize}
\item
  click on the \texttt{Knit} button in R Studio which by default will
  provide a HTML document in the RStudio viewer pane (the HTML will be
  stored in your working directory)
\item
  use pagedown's \texttt{chrome\_print} function in the YAML (uncomment
  line \#24 of this \texttt{Rmd} file) if additionally you want your
  HTML based web page to be printed to be PDF (the PDF will be stored in
  your working directory)
\item
  to ``live''-preview your pages do not click on the \texttt{Knit}
  button but use the \textbf{xaringan} (Xie 2021) RStudio add-in
  \emph{Infinite Moon Reader}. You can simply call the function
  \texttt{xaringan::inf\_mr()} (within your console). This will launch a
  local web server via the \textbf{servr} package (Xie 2021a) and
  display your pages in the RStudio viewer. Each time you save your
  document (\emph{Ctrl+S}) xaringan updates your pages in the viewer.
\item
  If you use the option \texttt{self\_contained:\ false} (see line \#21
  of this \texttt{Rmd} file) (change to true for a self-contained
  document, but it'll be a litte slower for Pandoc to render), don't
  click on the \texttt{Knit} button in RStudio. Use instead the
  \textbf{xaringan} (Xie 2021) RStudio add-in \emph{Infinite Moon
  Reader}.
\end{itemize}

\hypertarget{good-practices-for-reproducibility}{%
\section{Good practices for
reproducibility}\label{good-practices-for-reproducibility}}

Every researcher has his own optimized setup. Currently we would
recommend the following:

\begin{itemize}
\tightlist
\item
  Keep all files of your project (that matter for producing the PDF) in
  one folder without subfolders. You can zip and directly upload that
  folder to the \href{https://dataverse.harvard.edu/}{Harvard
  dataverse}.\footnote{Another good folder setup would be to store all
    files needed as input files for the R Markdown manuscript in a
    subfolder called ``input'' and all output files that are produced
    apart from paper.html and paper.pdf in a subfolder called
    ``output''.}
\item
  Make sure that filenames have a logic to them.

  \begin{itemize}
  \tightlist
  \item
    Main file with text/code: ``paper.rmd'', ``report.rmd''
  \item
    Data files: ``data\_xxxxxx.*''
  \item
    Image files: ``fig\_xxxxxx.*''
  \item
    Tables files: ``table\_xxxx.*''
  \item
    etc.
  \item
    Ideally, your filenames will correspond to the names in the paper.
    For instance, Figure 1 in the paper may have a corresponding file
    called \texttt{fig\_1\_xxxxx.pdf}.
  \end{itemize}
\item
  Use the document outline in R studio (Ctrl + Shift + O) when you work
  with R Markdown.
\item
  Name rchunks according to what they do or produce:

  \begin{itemize}
  \tightlist
  \item
    ``\texttt{fig-...}'' for chunks producing figures
  \item
    ``\texttt{table-...}'' for chunks producing tables
  \item
    ``\texttt{model-...}'' for chunks producing model estimates
  \item
    ``\texttt{import-...}'' for chunks importing data
  \item
    ``\texttt{recoding-...}'' for chunks in which data is recoded
  \end{itemize}
\item
  Use ``really'' informative variable names:

  \begin{itemize}
  \tightlist
  \item
    Q: What do you think does the variable \emph{trstep} measure? It
    actually measures trust in the European parliament.

    \begin{itemize}
    \tightlist
    \item
      How could we call this variable instead? Yes,
      \texttt{trust.european.parliament} which is longer but will
      probably be understood by another researcher in 50 years.
    \end{itemize}
  \item
    If your setup is truly reproducible you will probably re-use the
    variable names that you generate as variable names in the tables you
    produce. Hence, there is an incentive to use good names.
  \end{itemize}
\item
  Use unique identifiers in the final R Markdown document paper.rmd that
  you upload:

  \begin{itemize}
  \tightlist
  \item
    Think of someone who wants to produce Figure 1/Model 1/Table 1 in
    your paper but doesn't find it in your code\ldots{}

    \begin{itemize}
    \tightlist
    \item
      Name the chunks ``fig-1'', ``fig-2'' as the are named in the
      published paper.
    \item
      Name the chunks that produce tables ``table-1'', ``table-2'' etc.
      as they are named in the published paper.
    \item
      Name your statistical models in your R code ``M1'', ``M2'' as they
      are named in the published paper.
    \end{itemize}
  \end{itemize}
\end{itemize}

\hypertarget{additional-tricks-for-publishing}{%
\section{Additional tricks for
publishing}\label{additional-tricks-for-publishing}}

\begin{itemize}
\tightlist
\item
  Make your script anonymous

  \begin{itemize}
  \tightlist
  \item
    Simply put a \texttt{\textless{}!-\/-\ ...\ -\/-\textgreater{}}
    around any identifying information, e.g., author names, title
    footnote etc.
  \end{itemize}
\item
  Counting words

  \begin{itemize}
  \tightlist
  \item
    Use adobe acrobat (commerical software) to convert your file to a
    word file. Then open in word and delete all the parts that shouldn't
    go into the word count. The word count is displayed in the lower
    right.
  \item
    Use an one of the online services to count your words (search for
    ``pdf word count'')
  \end{itemize}
\item
  Appendix: You can change the numbering format for the appendix in the
  rmd file

  \begin{itemize}
  \tightlist
  \item
    What is still not possible in this document is to automatically have
    separate reference sections for paper and appendix.
  \end{itemize}
\item
  Journals may require you to use their tex style: Sometimes you can
  simply use their template in your rmarkdown file. See
  \href{https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/LDUMNY}{here}
  for a PLOS one example.
\end{itemize}

\hypertarget{citation-styles}{%
\section{Citation styles}\label{citation-styles}}

If your study needs to follow a particular citation style, you can set
the corresponding style in the header of your \texttt{.rmd} document. To
do so you have to download the corresponding \texttt{.csl} file.

In the present document we use the style of the American Sociological
Association and set it in the preamble with
\texttt{csl:\ american-sociological-association.csl}. However, you also
need to download the respective \texttt{.csl} file from the following
github page: https://github.com/citation-style-language/styles and copy
it into your working directory for it to work.

The github directory contains a wide variety of citation style files
depending on what discipline you work in.

\newpage

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-Bauer2018-dl}{}}%
Bauer, Paul. 2018. {``Writing a Reproducible Paper in {R} Markdown.''}
\emph{Open Science Framework Preprint}, December, 1--14.

\leavevmode\vadjust pre{\hypertarget{ref-Kirsop2005-ro}{}}%
Kirsop, Barbara, and Leslie Chan. 2005. {``Transforming Access to
Research Literature for Developing Countries.''} \emph{Serials Review}
31 (4): 246--55.

\leavevmode\vadjust pre{\hypertarget{ref-R2017}{}}%
R Core Team. 2017. \emph{R: A Language and Environment for Statistical
Computing}. Vienna, Austria: R Foundation for Statistical Computing.
\url{https://www.R-project.org/}.

\leavevmode\vadjust pre{\hypertarget{ref-Rstudio2015}{}}%
RStudio Team. 2015. \emph{RStudio: Integrated Development Environment
for r}. Boston, MA: RStudio, Inc. \url{http://www.rstudio.com/}.

\leavevmode\vadjust pre{\hypertarget{ref-knitr3}{}}%
Xie, Yihui. 2014. {``Knitr: A Comprehensive Tool for Reproducible
Research in {R}.''} In \emph{Implementing Reproducible Computational
Research}, edited by Victoria Stodden, Friedrich Leisch, and Roger D.
Peng. Chapman; Hall/CRC.
\url{http://www.crcpress.com/product/isbn/9781466561595}.

\leavevmode\vadjust pre{\hypertarget{ref-knitr2}{}}%
---------. 2015. \emph{Dynamic Documents with {R} and Knitr}. 2nd ed.
Boca Raton, Florida: Chapman; Hall/CRC. \url{https://yihui.name/knitr/}.

\leavevmode\vadjust pre{\hypertarget{ref-bookdown2}{}}%
---------. 2016. \emph{Bookdown: Authoring Books and Technical Documents
with {R} Markdown}. Boca Raton, Florida: Chapman; Hall/CRC.
\url{https://github.com/rstudio/bookdown}.

\leavevmode\vadjust pre{\hypertarget{ref-bookdown1}{}}%
---------. 2017. \emph{Bookdown: Authoring Books and Technical Documents
with r Markdown}. \url{https://github.com/rstudio/bookdown}.

\leavevmode\vadjust pre{\hypertarget{ref-knitr1}{}}%
---------. 2018. \emph{Knitr: A General-Purpose Package for Dynamic
Report Generation in r}. \url{https://yihui.name/knitr/}.

\leavevmode\vadjust pre{\hypertarget{ref-xaringan}{}}%
---------. 2021. \emph{Xaringan: Presentation Ninja}.
\url{https://CRAN.R-project.org/package=xaringan}.

\leavevmode\vadjust pre{\hypertarget{ref-Xie2021-bi}{}}%
Xie, Yihui, and Romain Lesur. 2021. {``Pagedown: Create Paged {HTML}
Documents for Printing from {R} Markdown.''}
\url{https://rstudio.github.io/pagedown/}.

\leavevmode\vadjust pre{\hypertarget{ref-Xie2021-ls}{}}%
Xie, Yihui, Romain Lesur, Brent Thorne, and Xianying Tan. 2021.
\emph{Pagedown: Paginate the HTML Output of r Markdown with CSS for
Print}. \url{https://CRAN.R-project.org/package=pagedown}.

\leavevmode\vadjust pre{\hypertarget{ref-kableextra}{}}%
Zhu, Hao. 2017. \emph{kableExtra: Construct Complex Table with 'Kable'
and Pipe Syntax}. \url{https://CRAN.R-project.org/package=kableExtra}.

\end{CSLReferences}

\newpage

\hypertarget{online-appendix}{%
\section*{Online appendix}\label{online-appendix}}
\addcontentsline{toc}{section}{Online appendix}

\hypertarget{sec:rsessioninfo}{%
\subsection{Attach R session info in appendix}\label{sec:rsessioninfo}}

Since R and R packages are constantly evolving you might want to add the
R session info that contains information on the R version as well as the
packages that are loaded.

\begin{verbatim}
R version 4.2.2 (2022-10-31 ucrt)
Platform: x86_64-w64-mingw32/x64 (64-bit)
Running under: Windows 10 x64 (build 17763)

Matrix products: default

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base

other attached packages:
[1] ggplot2_3.4.1      gt_0.8.0           dplyr_1.1.0        gdtools_0.3.0     
[5] flextable_0.8.5    modelsummary_1.3.0 kableExtra_1.3.4   knitr_1.42

loaded via a namespace (and not attached):
 [1] Rcpp_1.0.10        svglite_2.1.1      digest_0.6.31      utf8_1.2.3        
 [5] mime_0.12          R6_2.5.1           backports_1.4.1    evaluate_0.20     
 [9] httr_1.4.4         highr_0.10         pillar_1.8.1       rlang_1.0.6       
[13] curl_5.0.0         uuid_1.1-0         performance_0.10.2 rstudioapi_0.14   
[17] data.table_1.14.6  checkmate_2.1.0    rmarkdown_2.20     labeling_0.4.2    
[21] webshot_0.5.4      stringr_1.5.0      munsell_0.5.0      shiny_1.7.4       
[25] compiler_4.2.2     httpuv_1.6.9       xfun_0.37          parameters_0.20.2 
[29] pkgconfig_2.0.3    askpass_1.1        systemfonts_1.0.4  base64enc_0.1-3   
[33] gfonts_0.2.0       htmltools_0.5.4    insight_0.19.0     openssl_2.0.5     
[37] tidyselect_1.2.0   tibble_3.1.8       httpcode_0.3.0     fansi_1.0.4       
[41] viridisLite_0.4.1  withr_2.5.0        crayon_1.5.2       later_1.3.0       
[45] tables_0.9.10      crul_1.3           grid_4.2.2         gtable_0.3.1      
[49] jsonlite_1.8.4     xtable_1.8-4       lifecycle_1.0.3    bayestestR_0.13.0 
[53] magrittr_2.0.3     scales_1.2.1       datawizard_0.6.5   zip_2.2.2         
[57] cli_3.6.0          stringi_1.7.12     cachem_1.0.6       farver_2.1.1      
[61] promises_1.2.0.1   xml2_1.3.3         ellipsis_0.3.2     generics_0.1.3    
[65] vctrs_0.5.2        tools_4.2.2        glue_1.6.2         officer_0.5.2     
[69] fastmap_1.1.0      yaml_2.3.7         colorspace_2.1-0   rvest_1.0.3       
[73] memoise_2.0.1     
\end{verbatim}

\hypertarget{all-the-code-in-the-paper}{%
\subsection{All the code in the paper}\label{all-the-code-in-the-paper}}

To simply attach all the code you used in the PDF file in the appendix
see the R chunk in the underlying \texttt{.rmd} file:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{knitr}\SpecialCharTok{::}\NormalTok{opts\_chunk}\SpecialCharTok{$}\FunctionTok{set}\NormalTok{(}\AttributeTok{cache =} \ConstantTok{FALSE}\NormalTok{)}
\CommentTok{\# Use chache = TRUE if you want to speed up compilation}
\CommentTok{\# A function to allow for showing some of the inline code}
\NormalTok{rinline }\OtherTok{\textless{}{-}} \ControlFlowTok{function}\NormalTok{(code)\{}
\NormalTok{  html }\OtherTok{\textless{}{-}} \StringTok{\textquotesingle{}\textless{}code  class="r"\textgreater{}\textasciigrave{}\textasciigrave{}\textasciigrave{} \textasciigrave{}r CODE\textasciigrave{} \textasciigrave{}\textasciigrave{}\textasciigrave{}\textless{}/code\textgreater{}\textquotesingle{}}
  \FunctionTok{sub}\NormalTok{(}\StringTok{"CODE"}\NormalTok{, code, html)}
\NormalTok{\}}
\NormalTok{remotes}\SpecialCharTok{::}\FunctionTok{install\_github}\NormalTok{(}\StringTok{\textquotesingle{}rstudio/pagedown\textquotesingle{}}\NormalTok{)}
\FunctionTok{install.packages}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\StringTok{"rmarkdown"}\NormalTok{, }\StringTok{"knitr"}\NormalTok{, }\StringTok{"kableExtra"}\NormalTok{,}
                   \StringTok{"stargazer"}\NormalTok{, }\StringTok{"modelsummary"}\NormalTok{, }\StringTok{"knitr"}\NormalTok{, }\StringTok{"gt"}\NormalTok{))}
\FunctionTok{cat}\NormalTok{(}\FunctionTok{paste}\NormalTok{(}\StringTok{"\#"}\NormalTok{, }\FunctionTok{capture.output}\NormalTok{(}\FunctionTok{sessionInfo}\NormalTok{()), }\StringTok{"}\SpecialCharTok{\textbackslash{}n}\StringTok{"}\NormalTok{, }\AttributeTok{collapse =}\StringTok{""}\NormalTok{)) }
  \CommentTok{\# or use message() instead of cat()}
\NormalTok{x }\OtherTok{\textless{}{-}} \DecValTok{1}\SpecialCharTok{:}\DecValTok{10}
\NormalTok{x}
\NormalTok{data }\OtherTok{\textless{}{-}} \FunctionTok{read.csv}\NormalTok{(}\StringTok{"data.csv"}\NormalTok{)}
\FunctionTok{head}\NormalTok{(data)}
\FunctionTok{dput}\NormalTok{(data[}\DecValTok{1}\SpecialCharTok{:}\DecValTok{5}\NormalTok{,]) }\CommentTok{\# here we only take a subset}
\NormalTok{data }\OtherTok{\textless{}{-}} \FunctionTok{structure}\NormalTok{(}\FunctionTok{list}\NormalTok{(}\AttributeTok{Fertility =} \FunctionTok{c}\NormalTok{(}\FloatTok{80.2}\NormalTok{, }\FloatTok{83.1}\NormalTok{, }\FloatTok{92.5}\NormalTok{, }\FloatTok{85.8}\NormalTok{, }\FloatTok{76.9}\NormalTok{), }\AttributeTok{Agriculture =} \FunctionTok{c}\NormalTok{(}\DecValTok{17}\NormalTok{, }
\FloatTok{45.1}\NormalTok{, }\FloatTok{39.7}\NormalTok{, }\FloatTok{36.5}\NormalTok{, }\FloatTok{43.5}\NormalTok{), }\AttributeTok{Examination =} \FunctionTok{c}\NormalTok{(15L, 6L, 5L, 12L, 17L}
\NormalTok{), }\AttributeTok{Education =} \FunctionTok{c}\NormalTok{(12L, 9L, 5L, 7L, 15L), }\AttributeTok{Catholic =} \FunctionTok{c}\NormalTok{(}\FloatTok{9.96}\NormalTok{, }\FloatTok{84.84}\NormalTok{, }
\FloatTok{93.4}\NormalTok{, }\FloatTok{33.77}\NormalTok{, }\FloatTok{5.16}\NormalTok{), }\AttributeTok{Infant.Mortality =} \FunctionTok{c}\NormalTok{(}\FloatTok{22.2}\NormalTok{, }\FloatTok{22.2}\NormalTok{, }\FloatTok{20.2}\NormalTok{, }\FloatTok{20.3}\NormalTok{, }
\FloatTok{20.6}\NormalTok{)), }\AttributeTok{class =} \StringTok{"data.frame"}\NormalTok{, }\AttributeTok{row.names =} \FunctionTok{c}\NormalTok{(}\ConstantTok{NA}\NormalTok{, }\SpecialCharTok{{-}}\NormalTok{5L))}
\FunctionTok{library}\NormalTok{(knitr)}
\FunctionTok{library}\NormalTok{(kableExtra)}

\FunctionTok{kable}\NormalTok{(swiss[}\DecValTok{1}\SpecialCharTok{:}\DecValTok{10}\NormalTok{,], }\AttributeTok{row.names =} \ConstantTok{TRUE}\NormalTok{, }
      \AttributeTok{caption =} \StringTok{\textquotesingle{}Table with kable() and kablestyling()\textquotesingle{}}\NormalTok{, }
      \AttributeTok{format =} \StringTok{"html"}\NormalTok{, }\AttributeTok{booktabs =}\NormalTok{ T) }\SpecialCharTok{\%\textgreater{}\%}
        \FunctionTok{kable\_styling}\NormalTok{(}\AttributeTok{full\_width =}\NormalTok{ T, }
                      \AttributeTok{latex\_options =} \FunctionTok{c}\NormalTok{(}\StringTok{"striped"}\NormalTok{, }
                                        \StringTok{"scale\_down"}\NormalTok{,}
                                        \StringTok{"HOLD\_position"}\NormalTok{),}
                      \AttributeTok{font\_size =} \DecValTok{10}\NormalTok{)}

\FunctionTok{library}\NormalTok{(modelsummary)}
\FunctionTok{datasummary\_skim}\NormalTok{(swiss, }
                 \AttributeTok{type=}\StringTok{"numeric"}\NormalTok{, }
                 \AttributeTok{histogram=}\NormalTok{T, }
                 \AttributeTok{title =} \StringTok{"Summary: Numeric variables"}\NormalTok{)}
\CommentTok{\# Create categorical variables}
\NormalTok{swiss}\SpecialCharTok{$}\NormalTok{Education\_cat }\OtherTok{\textless{}{-}} \FunctionTok{cut}\NormalTok{(swiss}\SpecialCharTok{$}\NormalTok{Education, }
                   \AttributeTok{breaks=}\FunctionTok{c}\NormalTok{(}\SpecialCharTok{{-}}\ConstantTok{Inf}\NormalTok{, }\DecValTok{6}\NormalTok{, }\DecValTok{12}\NormalTok{, }\ConstantTok{Inf}\NormalTok{), }
                   \AttributeTok{labels=}\FunctionTok{c}\NormalTok{(}\StringTok{"low"}\NormalTok{,}\StringTok{"middle"}\NormalTok{,}\StringTok{"high"}\NormalTok{))}
\NormalTok{swiss}\SpecialCharTok{$}\NormalTok{Infant.Mortality\_cat }\OtherTok{\textless{}{-}} \FunctionTok{cut}\NormalTok{(swiss}\SpecialCharTok{$}\NormalTok{Infant.Mortality, }
                   \AttributeTok{breaks=}\FunctionTok{c}\NormalTok{(}\SpecialCharTok{{-}}\ConstantTok{Inf}\NormalTok{, }\FloatTok{18.15}\NormalTok{, }\FloatTok{21.70}\NormalTok{, }\ConstantTok{Inf}\NormalTok{), }
                   \AttributeTok{labels=}\FunctionTok{c}\NormalTok{(}\StringTok{"low"}\NormalTok{,}\StringTok{"middle"}\NormalTok{,}\StringTok{"high"}\NormalTok{))}

\FunctionTok{library}\NormalTok{(flextable)}
\NormalTok{tab\_cat }\OtherTok{\textless{}{-}} \FunctionTok{datasummary\_skim}\NormalTok{(swiss, }
                            \AttributeTok{type=}\StringTok{"categorical"}\NormalTok{, }
                            \AttributeTok{title =} \StringTok{"Summary: Categorical variables"}\NormalTok{,}
                            \AttributeTok{output =} \StringTok{\textquotesingle{}flextable\textquotesingle{}}\NormalTok{)}

\CommentTok{\# additionally we want to change the font, fontsize and spacing}
\FunctionTok{library}\NormalTok{(}\StringTok{"gdtools"}\NormalTok{)}

\FunctionTok{library}\NormalTok{(dplyr)}
\NormalTok{tab\_cat }\OtherTok{\textless{}{-}}\NormalTok{ tab\_cat }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{font}\NormalTok{(}\AttributeTok{fontname=}\StringTok{"Times New Roman"}\NormalTok{, }\AttributeTok{part=}\StringTok{"header"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{font}\NormalTok{(}\AttributeTok{fontname=}\StringTok{"Times New Roman"}\NormalTok{, }\AttributeTok{j=}\DecValTok{1}\SpecialCharTok{:}\DecValTok{4}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{fontsize}\NormalTok{(}\AttributeTok{size=}\DecValTok{12}\NormalTok{, }\AttributeTok{part=}\StringTok{"header"}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{fontsize}\NormalTok{(}\AttributeTok{size=}\DecValTok{10}\NormalTok{, }\AttributeTok{j=}\DecValTok{1}\SpecialCharTok{:}\DecValTok{4}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%} 
  \FunctionTok{line\_spacing}\NormalTok{(}\AttributeTok{space =} \FloatTok{0.3}\NormalTok{, }\AttributeTok{part =} \StringTok{"all"}\NormalTok{)}

\NormalTok{tab\_cat}
\FunctionTok{library}\NormalTok{(modelsummary)}
\NormalTok{M1 }\OtherTok{\textless{}{-}} \FunctionTok{lm}\NormalTok{(Fertility }\SpecialCharTok{\textasciitilde{}}\NormalTok{ Education }\SpecialCharTok{+}\NormalTok{ Agriculture, }\AttributeTok{data =}\NormalTok{ swiss)}
\NormalTok{M2 }\OtherTok{\textless{}{-}} \FunctionTok{lm}\NormalTok{(Fertility }\SpecialCharTok{\textasciitilde{}}\NormalTok{ Education }\SpecialCharTok{+}\NormalTok{ Catholic, }\AttributeTok{data =}\NormalTok{ swiss)}
\NormalTok{M3 }\OtherTok{\textless{}{-}} \FunctionTok{lm}\NormalTok{(Fertility }\SpecialCharTok{\textasciitilde{}}\NormalTok{ Education }\SpecialCharTok{+}\NormalTok{ Infant.Mortality }\SpecialCharTok{+}\NormalTok{ Agriculture, }\AttributeTok{data =}\NormalTok{ swiss)}
\NormalTok{models }\OtherTok{\textless{}{-}} \FunctionTok{list}\NormalTok{(}\StringTok{"M1"} \OtherTok{=}\NormalTok{ M1, }\StringTok{"M2"} \OtherTok{=}\NormalTok{  M2, }\StringTok{"M3"} \OtherTok{=}\NormalTok{ M3)}

\FunctionTok{library}\NormalTok{(gt)}
\CommentTok{\# additionally we want to change the font, font size and spacing}
\FunctionTok{modelsummary}\NormalTok{(models,}
             \AttributeTok{title =} \StringTok{\textquotesingle{}Linear regression\textquotesingle{}}\NormalTok{,}
             \AttributeTok{output =} \StringTok{\textquotesingle{}gt\textquotesingle{}}\NormalTok{,}
             \AttributeTok{notes =} \StringTok{"Notes: some notes..."}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
    \FunctionTok{tab\_spanner}\NormalTok{(}\AttributeTok{label =} \StringTok{\textquotesingle{}Dependent variable: Fertility\textquotesingle{}}\NormalTok{, }\AttributeTok{columns =} \DecValTok{2}\SpecialCharTok{:}\DecValTok{4}\NormalTok{) }\SpecialCharTok{\%\textgreater{}\%}
  \FunctionTok{tab\_options}\NormalTok{(}
    \AttributeTok{table.font.size =} \DecValTok{10}\NormalTok{,}
    \AttributeTok{data\_row.padding =} \FunctionTok{px}\NormalTok{(}\DecValTok{1}\NormalTok{),}
    \AttributeTok{table.border.top.color =} \StringTok{"white"}\NormalTok{,}
    \AttributeTok{heading.border.bottom.color =} \StringTok{"black"}\NormalTok{,}
    \AttributeTok{row\_group.border.top.color =} \StringTok{"black"}\NormalTok{,}
    \AttributeTok{row\_group.border.bottom.color =} \StringTok{"white"}\NormalTok{,}
    \AttributeTok{table.border.bottom.color =} \StringTok{"white"}\NormalTok{,}
    \AttributeTok{column\_labels.border.top.color =} \StringTok{"black"}\NormalTok{,}
    \AttributeTok{column\_labels.border.bottom.color =} \StringTok{"black"}\NormalTok{,}
    \AttributeTok{table\_body.border.bottom.color =} \StringTok{"black"}\NormalTok{,}
    \AttributeTok{table\_body.hlines.color =} \StringTok{"white"}
\NormalTok{  )}

\FunctionTok{plot}\NormalTok{(swiss}\SpecialCharTok{$}\NormalTok{Catholic, swiss}\SpecialCharTok{$}\NormalTok{Fertility)}
\FunctionTok{library}\NormalTok{(ggplot2)}
\FunctionTok{ggplot}\NormalTok{(swiss, }\FunctionTok{aes}\NormalTok{(}\AttributeTok{x=}\NormalTok{Catholic, }\AttributeTok{y=}\NormalTok{Fertility, }\AttributeTok{shape=}\NormalTok{Education\_cat)) }\SpecialCharTok{+} \FunctionTok{geom\_point}\NormalTok{() }\SpecialCharTok{+}
  \FunctionTok{labs}\NormalTok{(}\AttributeTok{x=}\StringTok{"Agriculture"}\NormalTok{, }\AttributeTok{y =} \StringTok{"Fertility"}\NormalTok{, }
       \AttributeTok{shape=}\StringTok{"Education"}\NormalTok{) }\SpecialCharTok{+} \FunctionTok{theme\_classic}\NormalTok{()}
\FunctionTok{print}\NormalTok{(}\FunctionTok{sessionInfo}\NormalTok{(), }\AttributeTok{local =} \ConstantTok{FALSE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\end{document}

### 2023-02-17 paper.qmd
Update abstract

@@
 date: 'First version: 16 February, 2023<br>This version: `r format(Sys.time(), "%d %B, %Y")`'
-abstract: 'The present paper provides a template to write a reproducible scientific paper with R Markdown and Pagedown.^[Based on an earlier R Markdown template that uses Latex and can be downloaded under https://osf.io/q395s [see @Bauer2018-dl].] Below we outline some of the "tricks"/code (e.g., referencing tables, sections etc.) we had to figure out to produce this document.
-The underlying files which produce this document can be [downloaded here](https://drive.google.com/drive/folders/1rWzj-Bu1EKqkSuE1gaFzduBJhzlThpkw?usp=sharing).
-Importantly, we also provide different CSS and HTML files that can be used to achieve a pdf output with the look of a "working paper".
-We are convinced that in the future there will be many improvements and developments with regards to RStudio, R markdown and Pagedown.
+abstract: 'The present paper provides a template to write a reproducible scientific paper with Quarto.
+It is based on earlier Pagedown/RMarkdown templates that can be downloaded [here](https://osf.io/k8jhx/) [@Bauer2018-dl] and [here](https://osf.io/q395s) [@Bauer2021-dl].
+Below we outline some of the "tricks"/code (e.g., referencing tables, sections etc.) we had to figure out to produce this document.
+The underlying files which produce this document can be [downloaded here](https://drive.google.com/drive/folders/1eT_9WaineZRB-nw-cQhNOy16sWhd_t2s?usp=sharing).
+We are convinced that in the future there will be many improvements and developments with regards to RStudio and Quarto.
 We intend to update this file when we discover more convenient code.

### 2023-02-17 paper.qmd
Update some cross-references.

@@
 To see how referencing works simply see the different examples for figures, tables and sections below.
-For instance in Section \@ref(sec:tables) you can find different ways of referencing tables.
-The code of the underlying `paper.qmd` will show you how I referenced Section \@ref(sec:tables) right here namely with '`Section \@ref(sec:tables)`'.
+For instance in @sec-tables you can find different ways of referencing tables.
+The code of the underlying `paper.qmd` will show you how I referenced @sec-tables right here namely with '`@sec-tables`' whereas the corresponding section title was assigned the corresponding label '`# Tables {#sec-tables}`'.
 
@@
 
-# Tables {#sec:tables}
+# Tables {#sec-tables}
 
@@
 @tbl-2 provides an example.
-To reference the table produced by the chunk you need to add `#| label: tbl-x` at the start of the chunk, i.e., `#| label: tbl-2` and would reference it by adding "`@tbl-2`" in your text.
+To reference the table produced by the chunk you need to add `#| label: tbl-x` at the start of the chunk, i.e., `#| label: tbl-1` and would reference it by adding "`@tbl-1`" in your text.
+See below for the full chunk code (the one that actually produces the table is hidden).
 
- {{
+```{{r}}
+#| label: tbl-1
+#| tbl-cap: "Summary: Numeric variables"
 
-```{r table-2, echo=TRUE}
-#| label: tbl-2
+library(knitr)
+library(kableExtra)
+
+kable(swiss[1:5,], row.names = TRUE, 
+      caption = 'Table with kable() and kablestyling()', 
+      format = "html", booktabs = T) %>%
+        kable_styling(full_width = T, 
+                      latex_options = c("striped", 
+                                        "scale_down",
+                                        "HOLD_position"),
+                      font_size = 11)
+```
+
+```{r table-2, echo=FALSE}
+#| label: tbl-1
 #| tbl-cap: "Summary: Numeric variables"
@@
 
-@tbl-3 shows a summary table for numeric variables.
+@tbl-2 shows a summary table for numeric variables.
 
 ```{r table-3, echo=TRUE}
-#| label: tbl-3
+#| label: tbl-2
 #| tbl-cap: "Summary: Numeric variables"
@@
 
-@tbl-4 shows a summary table for categorical variables.
+@tbl-3 shows a summary table for categorical variables.
 
 ```{r table-4, echo=TRUE}
-#| label: tbl-4
+#| label: tbl-3
 #| tbl-cap: "Summary: Categorical variables"
@@
 
-@tbl-5 shows the output for a regression table. Make sure you name all the models you estimate (even if its 50) and explicitly refer to model names (M1, M2 etc.) in the text.
+@tbl-4 shows the output for a regression table. Make sure you name all the models you estimate (even if its 50) and explicitly refer to model names (M1, M2 etc.) in the text.
 
 ```{r}
-#| label: tbl-5
+#| label: tbl-4
 #| tbl-cap: "Linear regression"

### 2023-02-17 paper.qmd
Update author and date.

@@
 title: "Writing a reproducible paper with Quarto"
-author: "Paul C. Bauer, Camille Landesvatter"
+author:
+  - name: Paul C. Bauer
+    orcid: 0000-0002-8382-9724
+    email: mail@paulcbauer.de
+    affiliations:
+      - name: Mannheim Centre for European Social Research
+  - name: Camille Landesvatter
+    orcid: 0000-0003-0156-5364
+    email: camille.landesvatter@uni-mannheim.de
+    affiliations:
+      - name: Mannheim Centre for European Social Research
 format: html
-date: 'First version: 16 February, 2023<br>This version: `r format(Sys.time(), "%d %B, %Y")`'
+date: 'last-modified'
+date-format: "[First version:February  16, 2023<br>This version:] MMMM D, YYYY"
 abstract: 'The present paper provides a template to write a reproducible scientific paper with Quarto.

### 2023-03-03 paper.qmd
edits prerequisites and basics.

--- 2023-02-17 paper.qmd
@@
 ---
-title: "Writing a reproducible paper with Quarto"
-author: "Paul C. Bauer, Camille Landesvatter"
-format: html
-date: 'First version: 16 February, 2023<br>This version: `r format(Sys.time(), "%d %B, %Y")`'
+title: "Writing a reproducible paper with RStudio and Quarto"
+author:
+  - name: Paul C. Bauer
+    orcid: 0000-0002-8382-9724
+    email: mail@paulcbauer.de
+    affiliations:
+      - name: Mannheim Centre for European Social Research
+  - name: Camille Landesvatter
+    orcid: 0000-0003-0156-5364
+    email: camille.landesvatter@uni-mannheim.de
+    affiliations:
+      - name: Mannheim Centre for European Social Research
+format:
+  html: default
+  pdf: default
+date: 'last-modified'
+date-format: "[First version:February  16, 2023<br>This version:] MMMM D, YYYY"
 abstract: 'The present paper provides a template to write a reproducible scientific paper with Quarto.
@@
 
-# Prerequesites
+# Prerequisites
 
-We assume that you are using R on a day-to-day basis and you may have even started to work in R Markdown.
-If you don't know what R Markdown is there are many great resources that you should use (e.g.
+Quarto is a multi-language, next-generation version of R Markdown from RStudio.
+Multi-language means you can use R, Python, Julia and other languages to create quarto documents.
+In this template we’ll focus on showing you how to use **RStudio with Quarto**.
+
+To name prerequisites, we assume that you are familiar with using R and R Markdown.
+If you don't know what R Markdown is there are many great resources that you can use (e.g.
 watch this [short video](https://vimeo.com/178485416)).
-An older template [see @Bauer2018-dl; https://osf.io/q395s] on which this newer template is based, may provide a quick entry point to writing a reproducible with R Markdown and Latex.
+An older template of ours [see @Bauer2018-dl; https://osf.io/q395s] may provide a quick entry point to writing a reproducible manuscript with R Markdown and Latex.
+To further deepen your understanding of the different components underlying this template, go on and learn about R, [Markdown](https://en.wikipedia.org/wiki/Markdown), [R Markdown](https://rmarkdown.rstudio.com/lesson-1.html) and [Quarto](https://quarto.org/).
 
-Based on R Markdown, Pagedown allows you to create custom and well-formatted (paged) HTML Documents.
-For a comprehensive overview watch [this video](https://www.rstudio.com/resources/rstudioconf-2019/pagedown-creating-beautiful-pdfs-with-r-markdown-and-css/) which is a record of a talk introducing `pagedown` given by Yihui Xie (who in addition to Romain Lesur developed the `pagedown` package).
-If you are not in a video watching mood find the slides [here](https://slides.yihui.org/2019-rstudio-conf-pagedown.html#1).
+Like R Markdown, Quarto uses `Knitr` to execute R code (hence you should also be able to render existing Rmd files without modification), and can render documents of a dozens of formats (e.g., PDF, HTML).
+For a comprehensive overview watch [this video](https://www.youtube.com/watch?v=yvi5uXQMvu4&t=708s&ab_channel=PositPBC) which is a record of a talk introducing `quarto` given by Tom Mock from RStudio.
 
@@
 
-* ...install [R](https://www.r-project.org/) and [Rstudio](https://www.rstudio.com/) (most recent versions) [@R2017; @Rstudio2015].
-* ...install the "pagedown"-package from github using the code below [@Xie2021-bi; @Xie2021-ls].
+* ...install [R](https://www.r-project.org/) and [Rstudio](https://posit.co/download/rstudio-desktop/) (important: most recent versions, e.g. RStudio v2022.07) [@R2017; @Rstudio2015].
+* ...install the "rmarkdown"-package from CRAN using the code below [@markdown2017].
+```{r eval=FALSE, echo=TRUE}
+install.packages("rmarkdown")
+```
+* ...install [tinytex](https://yihui.name/tinytex/), a lightweight version of Tex Live [@markdown2017; @tinytex] in order to compile PDFs.
 ```{r eval=FALSE, include=TRUE}
-remotes::install_github('rstudio/pagedown')
+install.packages(c('tinytex', 'rmarkdown'))
+tinytex::install_tinytex()
 ```
 * ...also install the packages below using the code below [@bookdown1; @bookdown2; @knitr1; @knitr2; @knitr3; @kableextra].
-```{r eval=FALSE, include=TRUE}
-install.packages(c("rmarkdown", "knitr", "kableExtra",
+```{r eval=FALSE, echo=TRUE}
+install.packages(c("knitr", "kableExtra",
                    "stargazer", "modelsummary", "knitr", "gt"))
 ```
-* ...download the 4 input files we created --- `paper.qmd`, `references.bib`, `data.csv` and `american-sociological-association.csl` --- from [this folder](https://drive.google.com/drive/folders/1rWzj-Bu1EKqkSuE1gaFzduBJhzlThpkw?usp=sharing).
+* ...download the 3 input files we created --- `paper.qmd`, `references.bib` and `data.csv` --- from [this folder](https://drive.google.com/drive/folders/1rWzj-Bu1EKqkSuE1gaFzduBJhzlThpkw?usp=sharing).
 Ignore the other files.
-* ...also download the 4 styling files we created: `wp_paged.html`, `wp.css`, `wp-fonts.css` and `wp-pages.css`.
-* ...store all 8 files from above together in one folder (and use this folder as your working directory later on)
-* ...learn R and read about the other underlying components namely [Markdown](https://en.wikipedia.org/wiki/Markdown), [R Markdown](https://rmarkdown.rstudio.com/lesson-1.html) and [Pagedown](https://pagedown.rbind.io/).
-* ...pagedown comes with several Rmd-templates (presentations, poster, thesis, etc.) and via this review we provide another template for a working paper style.
-If however you want to modify single aspects or create your own template, you will need to at least gain some basic skills in [CSS](https://www.w3schools.com/css/) and [HTML](https://www.w3schools.com/html/).
+* ...finally store all files from above together in one folder (and use this folder as your working directory later on)
 
-# Basics: Input files, output files and the YAML header
+# Basics: Input files
 
-All the files you need to produce the present PDF file are:
-
-1. the input files:
+All the input files you need to produce the present PDF file are:
 
@@
 * `references.bib` (the bibliography).
-    + I use paperpile to manage my references and export the `.bib` file into the folder that contains my `.qmd` file.
+    + We use paperpile to manage references and export the `.bib` file into the folder that contains the `.qmd` file.
 * `data.csv` (some raw data).
-* `american-sociological-association.csl` (defines the style of your bibliography).^[You can download various citation style files from this webpage: https://github.com/citation-style-language/styles.]
-
-2. the "styling" files:
-
-Basically, these are files you will need to specify in the YAML of your rmd-file, so that R and ultimately pagedown recognizes the certain style you want to achieve for your document.
-With using our templates, you will create a document that has the "look" of a working paper (we based our files on the [jss_paged pagedown format](https://github.com/rstudio/pagedown#journal-of-statistical-software-article-pagedownjss_paged)).
-
-* `wp_paged.html` (based on `jss_paged.html`)
-* `wp.css`
-* `wp-fonts.css`
-* `wp-pages.css`
-
-Take `paper.qmd` (the underlying R Markdown file of this pdf) and have a look at the YAML (line \#18 - \#22) to see how to specifiy these files.
-Basically, what happens here is that within the [jss_paged function](https://rdrr.io/cran/pagedown/man/jss_paged.html) we additionally specify that we want to use custom CSS and custom HTML.
 
 [Download these files](https://drive.google.com/drive/folders/1rWzj-Bu1EKqkSuE1gaFzduBJhzlThpkw?usp=sharing) and save them into a folder.
-Close R/Rstudio and directly open `paper_pagedown.qmd` with RStudio.
+Close R/Rstudio and directly open `paper.qmd` with RStudio.
 Doing so assures that the working directory is set to the folder that contains `paper.qmd` and the other files.^[You can always check your working directory in R with `getwd()`.]
 
-Once you run/compile the `paper.qmd` file in Rstudio it creates a output file called `paper_pagedown.html`.
+# Basics: Output files and the YAML header
 
-By using pagedown's `chrome_print` function in the YAML (line \#25) your html based web page will be printed to `paper_pagedown.pdf` (the one you are reading right now).
+Quarto allows you to produce documents in various formats (e.g., HTML, PDF, MS Word).
+By default, once you run/render the `paper.qmd` file in Rstudio it creates an output file in HTML called `paper.html`.
+However in the YAML you can specify any alternative format (or formats) alongside different options.
 
-Both outputs will be saved in your working directory.
+Remember (cf. Section Prerequesites) that in order to compile PDFs you will need to install a recent distribution of LaTeX (e.g., TinyTex)
+
+If you want your document to be rendered in multiple output formats (e.g., HTML and PDF like we did here), you need to specify both outputs in your YAML header.
+Subsequently, you have two options on how to proceed for rendering:
+
+* first, use the Render Button in R Studio.
+If you click it without using the drop down menu, by default it will create HTML since it is the first format listed in our YAML.
+If you use the drop down menu of the "Render"-button, you can specify the desired output (e.g., HTML or PDF) but only if you specified it in your YAML.
+
+* second, if you are not using RStudio and/or you prefer to render from the R console, use the `quarto` package to render to all formats at the same time with the `quarto_render()` function:
+
+```{r eval=F, echo=T}
+quarto::quarto_render()
+```
+
+Beforehand don't forget to install the `quarto` package:
+
+```{r eval=F, echo=T}
+install.packages("quarto")
+```
+
+Here you can also again specify the name of the resulting document as well as the desired output format(s):
+
+```{r eval=F, echo=T}
+quarto::quarto_render(
+  "paper.qmd", 
+  output_format = c("pdf", "html", "docx")
+)
+
+#quarto::quarto_render("paper.qmd") # defaults to html
+```
+
+Your output(s) will be saved in your working directory.
+
+## Some more notes on the output format (your YAML header)
+
+Take `paper.qmd` (the underlying R Markdown file of this pdf) and have a look at the YAML (line \#14 - \#16) to see how to specify different outputs.
 

### 2023-03-03 paper.qmd
small edits and section "Why Quarto".

@@
 date-format: "[First version:February  16, 2023<br>This version:] MMMM D, YYYY"
-abstract: 'The present paper provides a template to write a reproducible scientific paper with Quarto.
-It is based on earlier Pagedown/RMarkdown templates that can be downloaded [here](https://osf.io/k8jhx/) [@Bauer2018-dl] and [here](https://osf.io/q395s) [@Bauer2021-dl].
+abstract: 'This paper provides a template to write a reproducible scientific paper with RStudio and Quarto.
+It falls into a series on other templates, namely on RMarkdown and pagedown that can be downloaded [here](https://osf.io/k8jhx/) [@Bauer2018-dl] and [here](https://osf.io/q395s) [@Bauer2021-dl].
 Below we outline some of the "tricks"/code (e.g., referencing tables, sections etc.) we had to figure out to produce this document.
@@
 
-Formatting text as PDF is probably one of the most widespread standards in the scientific community, especially when it comes to submitting papers and similar documents.
-The traditional way to well-formatted and good-looking PDFs is often through LaTeX or Word.
-However, if you have spent hours and hours debugging latex code (or getting it to run) you may be on the lookout for something new.
-
-The fairly new `pagedown` R package takes a completely new approach.
-While the main purpose of pagedown is to create high-quality PDFs, the idea is to take advantage of modern web technologies (HTML/JSS/Javascript) with which one can design web pages and eventually print those to PDF.
-
-While web pages are usually single-page scrollable documents, pagedown uses the JavaScript library `Paged.js` which allows documents to be paginated with elements like headers, footers and everything a readable scientific paper will need.
-Additionally, pagedown documents are based on R Markdown.
-In our view, Pagedown and the underlying technology may replace Latex in the long run.
-In the near future it should also be possible to produce a PDF with static graph and an equivalent html with interactive graps (see dicsussion [here](https://github.com/rstudio/pagedown/pull/87)).
+Quarto is an open-source, scientific and technical publishing system build on `pandoc`.
+Their creators from RStudio also call it [„the next generation of R Markdown“](https://quarto.org/docs/computations/r.html).
+Like RMarkdown, Quarto uses `Knitr` to execute `R` code, and is therefore able to render most existing `.rmd` files without having to make any adjustments.
+As in `rmd` and other more recent developments like R's `pagedown`-package, one of the main goals of Quarto is to weave together narrative text and code and produce nicely formatted output documents.
+Differences to RMarkdown are not only but especially due to two facts: first, Quarto is multi-lingual (R, Python, Julia, Observable JS, etc.) and second in addition to RMarkdown's workflow where the final conversion layer uses `pandoc` to create documents, Quarto uses pandoc with Lua filters (which include latex in pandoc metadata).
+This contributes to cross-language standardization of outputs.
+Put differently, it allows working regardless of language, since it allows shared syntax and shared format across languages and input formats.
 
@@
 * ...install [tinytex](https://yihui.name/tinytex/), a lightweight version of Tex Live [@markdown2017; @tinytex] in order to compile PDFs.
-```{r eval=FALSE, include=TRUE}
+```{r eval=FALSE, echo=TRUE}
 install.packages(c('tinytex', 'rmarkdown'))
@@
 Ignore the other files.
-* ...finally store all files from above together in one folder (and use this folder as your working directory later on)
+* ...finally store all files from above together in one folder (and use this folder as your working directory later on).
 
@@
 
-Remember (cf. Section Prerequesites) that in order to compile PDFs you will need to install a recent distribution of LaTeX (e.g., TinyTex)
+Remember (cf. Section Prerequesites) that in order to compile PDFs you will need to install a recent distribution of LaTeX (e.g., TinyTex).
 
 If you want your document to be rendered in multiple output formats (e.g., HTML and PDF like we did here), you need to specify both outputs in your YAML header.
+Take `paper.qmd` (the underlying quarto file of this pdf) and have a look at the YAML (line \#14 - \#16) to see how to specify different outputs.
+
 Subsequently, you have two options on how to proceed for rendering:
@@
 Your output(s) will be saved in your working directory.
-
-## Some more notes on the output format (your YAML header)
-
-Take `paper.qmd` (the underlying R Markdown file of this pdf) and have a look at the YAML (line \#14 - \#16) to see how to specify different outputs.
 

### 2023-03-03 paper.qmd
add link to quarto FAQ.
turn all html tables to eval=F so that pdf can be generated.

@@
 knitr::opts_chunk$set(cache = FALSE)
-# Use chache = TRUE if you want to speed up compilation
+# Use cache = TRUE if you want to speed up compilation
 # A function to allow for showing some of the inline code
@@
 
-Quarto is an open-source, scientific and technical publishing system build on `pandoc`.
+[Quarto](https://quarto.org/) is an open-source, scientific and technical publishing system build on `pandoc`.
 Their creators from RStudio also call it [„the next generation of R Markdown“](https://quarto.org/docs/computations/r.html).
@@
 To further deepen your understanding of the different components underlying this template, go on and learn about R, [Markdown](https://en.wikipedia.org/wiki/Markdown), [R Markdown](https://rmarkdown.rstudio.com/lesson-1.html) and [Quarto](https://quarto.org/).
+For Quarto also consider reading their [FAQ](https://quarto.org/docs/faq/rmarkdown.html).
 
@@
   "paper.qmd", 
-  output_format = c("pdf", "html", "docx")
+  output_format = c("pdf", "html")
 )
@@
 
-Producing good tables and referencing these tables within a R Markdown document has been a hassle but got much better. Examples that you may use are shown below.
+When producing a scientific Quarto document, you most likely cannot skip creating good tables and cross-referencing these tables throughout your document. Examples that you may use are shown below.
 
@@
 A great function is `kable()` (`knitr` package) in combination with `kableExtra`.
-@tbl-2 provides an example.
+@tbl-1 provides an example.
 To reference the table produced by the chunk you need to add `#| label: tbl-x` at the start of the chunk, i.e., `#| label: tbl-1` and would reference it by adding "`@tbl-1`" in your text.
@@
       caption = 'Table with kable() and kablestyling()',
-      format = "html", booktabs = T) %>%
+      booktabs = T, format = "html") %>%
         kable_styling(full_width = T, 
@@
                       font_size = 11)
+
 ```
 
-```{r table-2, echo=FALSE}
+```{r table-1, eval=F}
 #| label: tbl-1
@@
       caption = 'Table with kable() and kablestyling()', 
-      format = "html", booktabs = T) %>%
+      booktabs = T, format = "html") %>% 
         kable_styling(full_width = T, 
@@
 
-```{r table-3, echo=TRUE}
+```{r table-3, echo=TRUE, eval=F}
 #| label: tbl-2
@@
 
-```{r table-4, echo=TRUE}
+```{r table-4, echo=TRUE, eval=F}
 #| label: tbl-3
@@
 
-```{r}
+```{r eval=F}
 #| label: tbl-4
@@
 An example is to include estimates, e.g., the coefficient of `dist` of the model we ran above.
-`r rinline("round(coef(M1)[2], 2)")` will insert the coefficient as follows: `r round(coef(M1)[2], 2)`.
-Or `r rinline("3 + 7")` will insert a `r 3 + 7` in the text.
+`#r# rinline("round(coef(model1)[2], 2)")` will insert the coefficient as follows: `#r# round(coef(model1)[2], 2)`.
+Or `#r# rinline("3 + 7")` will insert a `r 3 + 7` in the text.
 Inline code/results that depend on earlier objects in your document will automatically be updated once you change those objects.
@@
 
-```{r}
+```{r eval=F}
 #| label: fig1
@@
 
-```{r}
+```{r eval=F}
 #| label: fig2

### 2023-03-03 paper.qmd
first edits tables (+ edits referencing section).

@@
 tbl-cap-location: top
+number-sections: true
 execute: 
@@
 ```{r eval=FALSE, echo=TRUE}
-install.packages(c('tinytex', 'rmarkdown'))
+install.packages('tinytex')
 tinytex::install_tinytex()
@@
 )
-
-#quarto::quarto_render("paper.qmd") # defaults to html
 ```
@@
 For instance in @sec-tables you can find different ways of referencing tables.
-The code of the underlying `paper.qmd` will show you how I referenced @sec-tables right here namely with '`@sec-tables`' whereas the corresponding section title was assigned the corresponding label '`# Tables {#sec-tables}`'.
+The code of the underlying `paper.qmd` will show you how we referenced @sec-tables right here namely with '`@sec-tables`' whereas the corresponding section title was assigned the corresponding label '`# Tables {#sec-tables}`'.
+Pay attention that when using section cross-references, you need to enable the number-sections option in your YAML (line \#22 of this `Rmd` file).
 
@@
 
-When producing a scientific Quarto document, you most likely cannot skip creating good tables and cross-referencing these tables throughout your document. Examples that you may use are shown below.
+When producing a Quarto document for scientific purposes, you most likely can't get around creating tables and cross-referencing these tables throughout your document.
+Examples that you could use are shown below.
+Importantly, we focus on tables that are customized for PDF output.
+Formatting text as PDF is probably one of the most widespread standards in the scientific community, especially when it comes to submitting papers and similar documents.
+However, where applicable we also mention libraries and code for html output.
 
@@
       caption = 'Table with kable() and kablestyling()',
-      booktabs = T, format = "html") %>%
+      booktabs = T, format="simple") %>%
         kable_styling(full_width = T, 
@@
 
-```{r table-1, eval=F}
+```{r}
 #| label: tbl-1
@@
       caption = 'Table with kable() and kablestyling()', 
-      booktabs = T, format = "html") %>% 
+      booktabs = T, format="simple") %>% 
         kable_styling(full_width = T, 
@@
 
+For HTML output simply change the format argument to "html".
+
+## Tables with stargazer(): Summary Statistics
+
+T@tbl-4 shows summary stats of your data using `stargazer()` [@hlavac2013stargazer] which offers extreme flexibility regarding table output (see `?stargazer`).
+Note how we added "results: 'asis'" to the chunk options.
+This is important in order to properly display the table.
+
+```{r}
+#| label: tbl-4
+#| tbl-cap: "Summary table with stargazer"
+#| results: "asis"
+
+library(stargazer)
+stargazer(swiss, 
+          title = "Summary table with stargazer",
+          label="tab1", 
+          table.placement = "H", 
+          #type="html",
+          header=FALSE)
+```
+
+If you want to create HTML output, definitely check out the modelsummary package.
+The `modelsummary` package provides a variety of tables and plots to summarize statistical models and data in R.
+Modellsummary plots and tables are highly customizable and they can be saved to almost all formats, e.g., HTML, PDF and Markdown.
+This makes it especially easy to embed them in dynamic documents.
+Please look at the package's extensive
+
 ## Tables with modelsummary
@@
 Modellsummary plots and tables are highly customizable and they can be saved to almost all formats, e.g., HTML, PDF and Markdown.
-This makes ist especially easy to embed them in dynamic documents.
+This makes it especially easy to embed them in dynamic documents.
 Please look at the package's extensive [documentation](https://vincentarelbundock.github.io/modelsummary/index.html) where they also show examples for almost any plot or table you might be looking for.
@@
 
-### Summarize numeric variables
+### Summarize numeric variables with modelsummary
 
@@
 
-```{r table-3, echo=TRUE, eval=F}
+```{r echo=TRUE, eval=F}
 #| label: tbl-2
@@
 
-### Summarize categorical variables
+### Summarize categorical variables with modelsummary
 
@@
 
-```{r table-4, echo=TRUE, eval=F}
+```{r echo=TRUE, eval=F}
 #| label: tbl-3
@@
 
-## Regression table
+## Regression Table with stargazer
 
-@tbl-4 shows the output for a regression table. Make sure you name all the models you estimate (even if its 50) and explicitly refer to model names (M1, M2 etc.) in the text.
+@tbl-6 shows the output for a regression table. Make sure you name all your models and explicitly refer to model names (M1, M2 etc.) in the text.
+
+```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, results="asis"}
+#| label: tbl-6
+#| tbl-cap: "Linear regression"
+
+library(stargazer)
+M1 <- lm(Fertility ~ Education + Agriculture, data = swiss)
+M2 <- lm(Fertility ~ Education + Catholic, data = swiss)
+M3 <- lm(Fertility ~ Education + Infant.Mortality + Agriculture, data = swiss)
+stargazer(M1, M2, M3,
+          title = "Regression table with stargazer",
+          label="tab2", 
+          table.placement = "H", 
+          column.labels = c("M1", "M2", "M3"),
+          model.numbers = FALSE,
+          header=FALSE)
+```
+
+For HTML output again consider using the modelsummary library:
+
+## Regression table with modelsummary
+
+@tbl-5 shows the output for a regression table. Make sure you name all the models you estimate (even if its 50) and explicitly refer to model names (M1, M2 etc.) in the text.
 
 ```{r eval=F}
-#| label: tbl-4
+#| label: tbl-5
 #| tbl-cap: "Linear regression"
@@
 An example is to include estimates, e.g., the coefficient of `dist` of the model we ran above.
-`#r# rinline("round(coef(model1)[2], 2)")` will insert the coefficient as follows: `#r# round(coef(model1)[2], 2)`.
-Or `#r# rinline("3 + 7")` will insert a `r 3 + 7` in the text.
+`r rinline("round(coef(M1)[2], 2)")` will insert the coefficient as follows: `r round(coef(M1)[2], 2)`.
+Or `r rinline("3 + 7")` will insert a `r 3 + 7` in the text.
 Inline code/results that depend on earlier objects in your document will automatically be updated once you change those objects.

### 2023-03-03 paper.qmd
adjust link to gdrive folder that contains alle files for download.

@@
 Below we outline some of the "tricks"/code (e.g., referencing tables, sections etc.) we had to figure out to produce this document.
-The underlying files which produce this document can be [downloaded here](https://drive.google.com/drive/folders/1eT_9WaineZRB-nw-cQhNOy16sWhd_t2s?usp=sharing).
+The underlying files which produce this document can be [downloaded here](https://drive.google.com/drive/u/0/folders/1eT_9WaineZRB-nw-cQhNOy16sWhd_t2s).
 We are convinced that in the future there will be many improvements and developments with regards to RStudio and Quarto.
@@
 ```
-* ...download the 3 input files we created --- `paper.qmd`, `references.bib` and `data.csv` --- from [this folder](https://drive.google.com/drive/folders/1rWzj-Bu1EKqkSuE1gaFzduBJhzlThpkw?usp=sharing).
+* ...download the 3 input files we created --- `paper.qmd`, `references.bib` and `data.csv` --- from [this folder](https://drive.google.com/drive/u/0/folders/1eT_9WaineZRB-nw-cQhNOy16sWhd_t2s).
 Ignore the other files.
@@
 
-[Download these files](https://drive.google.com/drive/folders/1rWzj-Bu1EKqkSuE1gaFzduBJhzlThpkw?usp=sharing) and save them into a folder.
+[Download these files](https://drive.google.com/drive/u/0/folders/1eT_9WaineZRB-nw-cQhNOy16sWhd_t2s) and save them into a folder.
 Close R/Rstudio and directly open `paper.qmd` with RStudio.

### 2023-03-05 paper.qmd
multiple edits.

@@
 bibliography: references.bib
+csl: american-sociological-association.csl
 tbl-cap-location: top
@@
 # Use cache = TRUE if you want to speed up compilation
+
 # A function to allow for showing some of the inline code
@@
 
-# Prerequisites
+# Prerequisites {#sec-prerequisites}
 
@@
 * ...install the "rmarkdown"-package from CRAN using the code below [@markdown2017].
-```{r eval=FALSE, echo=TRUE}
+
+```{{r}}
+install.packages("rmarkdown")
+```
+
+```{r eval=F}
 install.packages("rmarkdown")
@@
 * ...install [tinytex](https://yihui.name/tinytex/), a lightweight version of Tex Live [@markdown2017; @tinytex] in order to compile PDFs.
-```{r eval=FALSE, echo=TRUE}
+
+```{{r}}
 install.packages('tinytex')
@@
 ```
+
+```{r eval=F}
+install.packages('tinytex')
+tinytex::install_tinytex()
+```
+
 * ...also install the packages below using the code below [@bookdown1; @bookdown2; @knitr1; @knitr2; @knitr3; @kableextra].
-```{r eval=FALSE, echo=TRUE}
+
+```{{r}}
+install.packages(c("knitr", "kable", "kableExtra",
+                   "stargazer", "modelsummary", "gt"))
+```
+
+```{r eval=F}
 install.packages(c("knitr", "kableExtra",
@@
 
-# Basics: Output files and the YAML header
+# Basics: Output files and the YAML header {#sec-basics-2}
 
 Quarto allows you to produce documents in various formats (e.g., HTML, PDF, MS Word).
-By default, once you run/render the `paper.qmd` file in Rstudio it creates an output file in HTML called `paper.html`.
+By default, once you render the `paper.qmd` file in Rstudio it creates an output file in HTML called `paper.html`.
 However in the YAML you can specify any alternative format (or formats) alongside different options.
 
-Remember (cf. Section Prerequesites) that in order to compile PDFs you will need to install a recent distribution of LaTeX (e.g., TinyTex).
+Remember (see @sec-prerequisites) that in order to compile PDFs you will need to install a recent distribution of LaTeX (e.g., TinyTex).
 
@@
 
-Subsequently, you have two options on how to proceed for rendering:
-
-* first, use the Render Button in R Studio.
-If you click it without using the drop down menu, by default it will create HTML since it is the first format listed in our YAML.
-If you use the drop down menu of the "Render"-button, you can specify the desired output (e.g., HTML or PDF) but only if you specified it in your YAML.
-
-* second, if you are not using RStudio and/or you prefer to render from the R console, use the `quarto` package to render to all formats at the same time with the `quarto_render()` function:
-
-```{r eval=F, echo=T}
-quarto::quarto_render()
-```
-
-Beforehand don't forget to install the `quarto` package:
-
-```{r eval=F, echo=T}
-install.packages("quarto")
-```
-
-Here you can also again specify the name of the resulting document as well as the desired output format(s):
-
-```{r eval=F, echo=T}
-quarto::quarto_render(
-  "paper.qmd", 
-  output_format = c("pdf", "html")
-)
-```
-
-Your output(s) will be saved in your working directory.
+For instructions on how to render your final output document, please read @sec-compiling.
 
@@
 For instance in @sec-tables you can find different ways of referencing tables.
-The code of the underlying `paper.qmd` will show you how we referenced @sec-tables right here namely with '`@sec-tables`' whereas the corresponding section title was assigned the corresponding label '`# Tables {#sec-tables}`'.
-Pay attention that when using section cross-references, you need to enable the number-sections option in your YAML (line \#22 of this `Rmd` file).
+`paper.qmd` (the underlying quarto file of this PDF) will show you how we referenced @sec-tables right here namely with '`@sec-tables`' whereas the corresponding section title was assigned the corresponding label '`# Tables {#sec-tables}`'.
+Pay attention that when using section cross-references, you need to enable the number-sections option in your YAML (line \#22 of this `qmd` file).
 
@@
 
-I would recommend to use the command below and simply add it to the appendix as I did here in Appendix \@ref(sec:rsessioninfo).
+I would recommend to use the command below and simply add it to the appendix as we did here in Appendix @sec-rsessioninfo.
 This will make sure you always inform the reader about the package versions your relied on in your paper.
@@
 
-```{r fig-versioning, echo=TRUE, eval=FALSE}
+```{{r}}
+#|label: fig-versioning
+
+cat(paste("#", capture.output(sessionInfo()), "\n", collapse ="")) 
+  # or use message() instead of cat()
+```
+
+```{r eval=F}
+#|label: fig-versioning
+
 cat(paste("#", capture.output(sessionInfo()), "\n", collapse ="")) 
@@
 
+```{{r}}
+x <- 1:10
+x
+```
+
 ```{r}
@@
 
-Below we import an exemplary dataset (you can find `data.csv` in the folder with the other files).
+Below we import an exemplary dataset (you can find `data.csv` in the [folder](https://drive.google.com/drive/u/0/folders/1eT_9WaineZRB-nw-cQhNOy16sWhd_t2s) with the other files).
 
-```{r, echo=T, results="raw"}
+```{{r}}
+|echo=T
+|results="raw"
+data <- read.csv("data.csv")
+head(data)
+```
+
+```{r}
 data <- read.csv("data.csv")
@@
 
+```{{r}}
+dput(data[1:5,]) # here we only take a subset
+```
+
 ```{r}
@@
 
-```{r eval=FALSE, include=TRUE, echo=TRUE}
+```{{r}}
+data <- structure(list(Fertility = c(80.2, 83.1, 92.5, 85.8, 76.9), Agriculture = c(17, 
+45.1, 39.7, 36.5, 43.5), Examination = c(15L, 6L, 5L, 12L, 17L
+), Education = c(12L, 9L, 5L, 7L, 15L), Catholic = c(9.96, 84.84, 
+93.4, 33.77, 5.16), Infant.Mortality = c(22.2, 22.2, 20.2, 20.3, 
+20.6)), class = "data.frame", row.names = c(NA, -5L))
+```
+
+```{r}
 data <- structure(list(Fertility = c(80.2, 83.1, 92.5, 85.8, 76.9), Agriculture = c(17, 
@@
 
-When producing a Quarto document for scientific purposes, you most likely can't get around creating tables and cross-referencing these tables throughout your document.
-Examples that you could use are shown below.
+If you are creating a Quarto document for scientific purposes, you probably cannot avoid creating tables and cross-referencing those tables in your document.
+Below we show examples that use different R libraries.
 Importantly, we focus on tables that are customized for PDF output.
 Formatting text as PDF is probably one of the most widespread standards in the scientific community, especially when it comes to submitting papers and similar documents.
-However, where applicable we also mention libraries and code for html output.
+However, where applicable we also mention libraries and code for HTML output.
 
@@
 
-A great function is `kable()` (`knitr` package) in combination with `kableExtra`.
-@tbl-1 provides an example.
-To reference the table produced by the chunk you need to add `#| label: tbl-x` at the start of the chunk, i.e., `#| label: tbl-1` and would reference it by adding "`@tbl-1`" in your text.
-See below for the full chunk code (the one that actually produces the table is hidden).
+A great function is `kable()` from the `knitr` package in combination with `kableExtra`.
+@tbl-1 provides such an example.
+To reference the table produced by the chunk you need to add `#| label: tbl-x` at the start of the chunk, i.e., `#| label: tbl-1` so that you can reference it by adding "`@tbl-1`" in your text.
+See below for the full chunk code.
 
@@
 
-kable(swiss[1:5,], row.names = TRUE, 
+knitr::kable(swiss[1:5,], row.names = TRUE, 
       caption = 'Table with kable() and kablestyling()',
       booktabs = T, format="simple") %>%
-        kable_styling(full_width = T, 
+        kableExtra::kable_styling(full_width = T, 
                       latex_options = c("striped", 
@@
 
-kable(swiss[1:5,], row.names = TRUE, 
+knitr::kable(swiss[1:5,], row.names = TRUE, 
       caption = 'Table with kable() and kablestyling()', 
       booktabs = T, format="simple") %>% 
-        kable_styling(full_width = T, 
+        kableExtra::kable_styling(full_width = T, 
                       latex_options = c("striped", 
@@
 
-For HTML output simply change the format argument to "html".
+## Tables with stargazer()
 
-## Tables with stargazer(): Summary Statistics
+@tbl-2 was generated using the `stargazer()` R package [@hlavac2013stargazer] and shows summary statistics of our data.
 
-T@tbl-4 shows summary stats of your data using `stargazer()` [@hlavac2013stargazer] which offers extreme flexibility regarding table output (see `?stargazer`).
-Note how we added "results: 'asis'" to the chunk options.
-This is important in order to properly display the table.
-
-```{r}
-#| label: tbl-4
+```{{r}}
+#| label: tbl-2
 #| tbl-cap: "Summary table with stargazer"
@@
           title = "Summary table with stargazer",
-          label="tab1", 
           table.placement = "H", 
-          #type="html",
+          type="latex",
           header=FALSE)
@@
 
-If you want to create HTML output, definitely check out the modelsummary package.
-The `modelsummary` package provides a variety of tables and plots to summarize statistical models and data in R.
-Modellsummary plots and tables are highly customizable and they can be saved to almost all formats, e.g., HTML, PDF and Markdown.
-This makes it especially easy to embed them in dynamic documents.
-Please look at the package's extensive
+```{r}
+#| label: tbl-2
+#| tbl-cap: "Summary table with stargazer"
+#| results: "asis"
+
+library(stargazer)
+stargazer(swiss, 
+          title = "Summary table with stargazer",
+          label="tbl-2", 
+          table.placement = "H", 
+          type="latex",
+          header=FALSE)
+```
+
+In the above example we set the `type` argument to `latex` (default) which enables us to compile the table for PDF.
+This way, if you're reading the HTML output of this document, you won't be able to see the table (or it's cross-reference).
+This means, that dependent on your desired output format you need tp change the type argument.
+For example, use `type="html"` for HTML outfput formats.
+
+Lastly, note, how we specified `results: "asis"` into the code chunk options. This is important so that latex is properly printed to the PDF.
 
@@
 
-The `modelsummary` package provides a variety of tables and plots to summarize statistical models and data in R.
-Modellsummary plots and tables are highly customizable and they can be saved to almost all formats, e.g., HTML, PDF and Markdown.
+If you want to create well-formatted HTML output, consider learning about the `modelsummary` R package.
+`modelsummary` provides a variety of tables and plots to summarize statistical models and data in R.
+`modelsummary` plots and tables are highly customizable and they can be saved to almost all formats, e.g., HTML, PDF and Markdown.
 This makes it especially easy to embed them in dynamic documents.
@@
 
-@tbl-2 shows a summary table for numeric variables.
+@tbl-4 shows a summary table for numeric variables.
 
-```{r echo=TRUE, eval=F}
-#| label: tbl-2
+```{{r}}
+#| label: tbl-4
+#| tbl-cap: "Summary: Numeric variables"
+
+library(modelsummary)
+datasummary_skim(swiss, 
+                 type="numeric", 
+                 histogram=T, 
+                 title = "Summary: Numeric variables")
+```
+
+```{r eval=F}
+#| label: tbl-4
 #| tbl-cap: "Summary: Numeric variables"
@@
 
-@tbl-3 shows a summary table for categorical variables.
+@tbl-5 shows a summary table for categorical variables.
 
-```{r echo=TRUE, eval=F}
-#| label: tbl-3
+```{{r}}
+#| label: tbl-5
+#| tbl-cap: "Summary: Categorical variables"
+
+# Create categorical variables
+swiss$Education_cat <- cut(swiss$Education, 
+                   breaks=c(-Inf, 6, 12, Inf), 
+                   labels=c("low","middle","high"))
+swiss$Infant.Mortality_cat <- cut(swiss$Infant.Mortality, 
+                   breaks=c(-Inf, 18.15, 21.70, Inf), 
+                   labels=c("low","middle","high"))
+
+library(flextable)
+tab_cat <- datasummary_skim(swiss, 
+                            type="categorical", 
+                            output = 'flextable')
+
+# additionally we want to change the font, fontsize and spacing
+library("gdtools")
+
+library(dplyr)
+tab_cat <- tab_cat %>%
+  font(fontname="Times New Roman", part="header") %>%
+  font(fontname="Times New Roman", j=1:4) %>% 
+  fontsize(size=12, part="header") %>%
+  fontsize(size=10, j=1:4) %>% 
+  line_spacing(space = 0.3, part = "all")
+
+tab_cat
+```
+
+```{r eval=F}
+#| label: tbl-5
 #| tbl-cap: "Summary: Categorical variables"
@@
 
-For HTML output again consider using the modelsummary library:
-
 ## Regression table with modelsummary
 
-@tbl-5 shows the output for a regression table. Make sure you name all the models you estimate (even if its 50) and explicitly refer to model names (M1, M2 etc.) in the text.
+@tbl-7 shows the output for a regression table. Make sure you name all the models you estimate (even if its 50) and explicitly refer to model names (M1, M2 etc.) in the text.
 
 ```{r eval=F}
-#| label: tbl-5
+#| label: tbl-7
 #| tbl-cap: "Linear regression"
@@
 
-# Compiling the document
+# Compiling the document {#sec-compiling}
 
-To view your paper, pagedown requires a web server (since it is based on paged.js)^[open-source library to paginate content in the browser].
-By compiling a document, R Studio will display your HTML page through a local web server, i.e., paged.js will work in RStudio Viewer.
+Eventually you have two options on how to render/compile/knit your document:
 
-There are several options, depending on your intention:
+* first, use the Render Button in R Studio.
+If you click it without using the drop down menu, by default it will create HTML since it is the first format listed in our YAML.
+Always remember to specify your desired output formats in your YAML (as described in @sec-basics-2) If you use the drop down menu of the "Render"-button, you can specify the desired output (e.g., HTML or PDF) but only if you specified it in your YAML.
 
-- click on the `Knit` button in R Studio which by default will provide a HTML document in the RStudio viewer pane (the HTML will be stored in your working directory)
-- use pagedown's `chrome_print` function in the YAML (uncomment line \#24 of this `Rmd` file) if additionally you want your HTML based web page to be printed to be PDF (the PDF will be stored in your working directory)
-- to "live"-preview your pages do not click on the `Knit` button but use the **xaringan** [@xaringan] RStudio add-in _Infinite Moon Reader_.
-You can simply call the function `xaringan::inf_mr()` (within your console).
-This will launch a local web server via the **servr** package (Xie 2021a) and display your pages in the RStudio viewer.
-Each time you save your document (_Ctrl+S_) xaringan updates your pages in the viewer.
+* second, if you are not using RStudio and/or you prefer to render from the R console, use the `quarto` package to render to all formats at the same time with the `quarto_render()` function:
 
-- If you use the option `self_contained: false` (see line \#21 of this `Rmd` file) (change to true for a self-contained document, but it'll be a litte slower for Pandoc to render), don't click on the `Knit` button in RStudio.
-Use instead the **xaringan** [@xaringan] RStudio add-in _Infinite Moon Reader_.
+```{{r}}
+quarto::quarto_render()
+```
+
+```{r eval=F}
+quarto::quarto_render()
+```
+
+Beforehand don't forget to install the `quarto` package [@quarto]:
+
+```{{r}}
+install.packages("quarto")
+```
+
+```{r eval=F}
+install.packages("quarto")
+```
+
+Here you can also again specify the name of the resulting document as well as the desired output format(s):
+
+```{{r}}
+quarto::quarto_render(
+  "paper.qmd", 
+  output_format = c("pdf", "html")
+)
+```
+
+```{r eval=F}
+quarto::quarto_render(
+  "paper.qmd", 
+  output_format = c("pdf", "html")
+)
+```
+
+Your output(s) will be saved in your working directory.
 
@@
 
+Also read [quarto's official documentation](https://quarto.org/docs/authoring/footnotes-and-citations.html) about Citations and Footnotes.
+
 \newpage
@@
 
-## Attach R session info in appendix {#sec:rsessioninfo}
+## Attach R session info in appendix {#sec-rsessioninfo}
 

### 2023-03-05 paper.qmd
edits.

@@
 #|label: fig-versioning
-
 cat(paste("#", capture.output(sessionInfo()), "\n", collapse ="")) 
@@
 #| tbl-cap: "Summary: Numeric variables"
-
 library(knitr)
@@
                       font_size = 11)
-
 ```
@@
 #| results: "asis"
-
 library(stargazer)
@@
 #| tbl-cap: "Summary: Numeric variables"
-
 library(modelsummary)
@@
                  histogram=T, 
+                 output="default",
                  title = "Summary: Numeric variables")
@@
 #| tbl-cap: "Summary: Categorical variables"
-
 # Create categorical variables
@@
 
-```{r eval=F}
+```{r}
 #| label: tbl-5
 #| tbl-cap: "Summary: Categorical variables"
-
+#| results: "asis"
 # Create categorical variables
@@
 library(flextable)
+library(modelsummary)
 tab_cat <- datasummary_skim(swiss, 
@@
 #| tbl-cap: "Linear regression"
-
 library(stargazer)
@@
 #| tbl-cap: "Linear regression"
-
 library(modelsummary)
@@
   )
-
 ```
@@
 ## R base graphs
-Inserting figures can be slightly more complicated. Ideally, we would produce and insert them directly in the `.qmd` file. It's relatively simple to insert R base graphs as you can see in @fig1.
+Inserting figures can be slightly more complicated. Ideally, we would produce and insert them directly in the `.qmd` file. It's relatively simple to insert R base graphs as you can see in @fig-1.
 
-```{r eval=F}
-#| label: fig1
+```{{r}}
+#| label: fig-1
 #| fig-cap: "Scatterplot of Speed and Distance"
-
 plot(swiss$Catholic, swiss$Fertility)
@@
 
-But it turns out that it doesn't always work so well.
+```{r}
+#| label: fig-1
+#| fig-cap: "Scatterplot of Speed and Distance"
+plot(swiss$Catholic, swiss$Fertility)
+```
 
 ## ggplot2 graphs
-Same is true for ggplot2 as you can see in @fig2.
 
-```{r eval=F}
-#| label: fig2
+Below in @fig-2 we also show a ggplot2 example.
+
+```{{r}}
+#| label: fig-2
 #| fig-cap: "Miles per gallon according to the weight"
+library(ggplot2)
+plot <- ggplot(swiss, aes(x=Catholic, y=Fertility, shape=Education_cat)) +
+  geom_point() +
+  labs(x="Agriculture", y = "Fertility", shape="Education") +
+  theme_classic()
+plot
+```
 
+```{r}
+#| label: fig-2
+#| fig-cap: "Miles per gallon according to the weight"
 library(ggplot2)
-ggplot(swiss, aes(x=Catholic, y=Fertility, shape=Education_cat)) + geom_point() +
-  labs(x="Agriculture", y = "Fertility", 
-       shape="Education") + theme_classic()
+plot <- ggplot(swiss, aes(x=Catholic, y=Fertility, shape=Education_cat)) +
+  geom_point() +
+  labs(x="Agriculture", y = "Fertility", shape="Education") +
+  theme_classic()
+plot
 ```

### 2023-03-05 paper.qmd
edits.

@@
   html: default
-  pdf: default
+  pdf:
+    number-sections: true
+    colorlinks: true
+    keeptex: true
 date: 'last-modified'
-date-format: "[First version:February  16, 2023<br>This version:] MMMM D, YYYY"
+date-format: '[First version: February  16, 2023 <br>This version:] MMMM D, YYYY'
 abstract: 'This paper provides a template to write a reproducible scientific paper with RStudio and Quarto.
@@
 
-* **Errors**: Manual steps in data analysis (e.g., manually copy/pasting values into a table etc.) may introduce errors. R Markdown allows you to **automatize** such steps and/or avoid them.
+* **Errors**: Manual steps in data analysis (e.g., manually copy/pasting values into a table etc.) may introduce errors. Quarto allows you to **automatize** such steps and/or avoid them.
 
@@
 
-* ...also install the packages below using the code below [@bookdown1; @bookdown2; @knitr1; @knitr2; @knitr3; @kableextra].
+* ...also install the packages below using the code below [@knitr1; @knitr2; @knitr3; @kableextra; @hlavac2013stargazer; @modelsummary; @gt].
 
 ```{{r}}
-install.packages(c("knitr", "kable", "kableExtra",
+install.packages(c("knitr", "kableExtra",
                    "stargazer", "modelsummary", "gt"))
@@
 
-```{r echo=TRUE, message=FALSE, warning=FALSE, paged.print=FALSE, results="asis"}
+```{r echo=TRUE, paged.print=FALSE, results="asis"}
 #| label: tbl-6
@@
 ## R base graphs
-Inserting figures can be slightly more complicated. Ideally, we would produce and insert them directly in the `.qmd` file. It's relatively simple to insert R base graphs as you can see in @fig-1.
+A fully reproducible manuscript would create diagrams directly in the `.qmd` file and insert them in the appropriate place. It's relatively simple to insert R base graphs as you can see in @fig-1.
 
@@
 
-Below in @fig-2 we also show a ggplot2 example.
+Below in @fig-2 we also show a example witrh R's `ggplot2` package.
 
@@
 
-```{r}
+```{r fig-2, fig.align="center", fig.pos="H", fig.width=6, fig.height=3}
 #| label: fig-2
-#| fig-cap: "Miles per gallon according to the weight"
+#| fig-cap: "Fertility and Agriculture by Education"
 library(ggplot2)

### 2023-03-08 paper.qmd
make stargazer section only visible in PDF and not in HTML.

@@
 # Use cache = TRUE if you want to speed up compilation
+
+knitr::opts_knit$set(output.format = "html")  # Set to "html" for HTML output
 
@@
 
-## Tables with stargazer()
+<!-- the stargazer section is only shown when the output format is PDF and not HTML -->
 
-@tbl-2 was generated using the `stargazer()` R package [@hlavac2013stargazer] and shows summary statistics of our data.
+`r if (knitr::is_latex_output()) "## Tables with stargazer()" else ""`
 
-```{{r}}
-#| label: tbl-2
-#| tbl-cap: "Summary table with stargazer"
-#| results: "asis"
-library(stargazer)
-stargazer(swiss, 
-          title = "Summary table with stargazer",
-          table.placement = "H", 
-          type="latex",
-          header=FALSE)
-```
+`r if (knitr::is_latex_output()) "@tbl-2 was generated using the stargazer() R package [@hlavac2013stargazer] and shows summary statistics of our data." else ""`
 
-```{r}
+```{r, eval = knitr::is_latex_output()}
 #| label: tbl-2
@@
 
-In the above example we set the `type` argument to `latex` (default) which enables us to compile the table for PDF.
+`r if (knitr::is_latex_output()) "In the above example we set the 'type' argument to 'latex' (default) which enables us to compile the table for PDF.
 This way, if you're reading the HTML output of this document, you won't be able to see the table (or it's cross-reference).
-This means, that dependent on your desired output format you need tp change the type argument.
-For example, use `type="html"` for HTML outfput formats.
-
-Lastly, note, how we specified `results: "asis"` into the code chunk options. This is important so that latex is properly printed to the PDF.
+This means, that dependent on your desired output format you need to change the type argument.
+For example, use 'type=html' for HTML output formats.<br> Lastly, note, how we specified 'results: asis' into the code chunk options.
+This is important so that latex is properly printed to the PDF." else ""`
 

### 2023-03-08 paper.qmd
display tables only conditional for pdf or html.

@@
 
-`r if (knitr::is_latex_output()) "## Tables with stargazer()" else ""`
+## Tables with stargazer()
 
-`r if (knitr::is_latex_output()) "@tbl-2 was generated using the stargazer() R package [@hlavac2013stargazer] and shows summary statistics of our data." else ""`
+`r if (knitr::is_latex_output()) "@tbl-2-pdf was generated using the stargazer() R package [@hlavac2013stargazer] and shows summary statistics of our data." else ""`
 
-```{r, eval = knitr::is_latex_output()}
+`r if (knitr::is_html_output()) "@tbl-2-html was generated using the stargazer() R package [@hlavac2013stargazer] and shows summary statistics of our data." else ""`
+
+```{{r}}
 #| label: tbl-2
@@
 stargazer(swiss, 
-          title = "Summary table with stargazer",
-          label="tbl-2", 
+          table.placement = "H", 
+          type="latex", # adjust to html for HTML output
+          header=FALSE)
+```
+
+```{r, eval = knitr::is_latex_output()}
+#| label: tbl-2-pdf
+#| tbl-cap: "Summary table with stargazer"
+#| results: "asis"
+library(stargazer)
+stargazer(swiss, 
           table.placement = "H", 
@@
 
-`r if (knitr::is_latex_output()) "In the above example we set the 'type' argument to 'latex' (default) which enables us to compile the table for PDF.
-This way, if you're reading the HTML output of this document, you won't be able to see the table (or it's cross-reference).
-This means, that dependent on your desired output format you need to change the type argument.
-For example, use 'type=html' for HTML output formats.<br> Lastly, note, how we specified 'results: asis' into the code chunk options.
-This is important so that latex is properly printed to the PDF." else ""`
+```{r eval = knitr::is_html_output()}
+#| label: tbl-2-html
+#| tbl-cap: "Summary table with stargazer"
+#| results: "asis"
+library(stargazer)
+stargazer(swiss,
+          table.placement = "H", 
+          type="html",
+          header=FALSE)
+```
+
+In the above example, note that the `type` argument defaults to `latex` and you have to insert `html` if you desire HTML output.
+Lastly, note, how we specified 'results: asis' into the code chunk options.
+This is important so that latex is properly printed to the PDF.
 
@@
 
-@tbl-4 shows a summary table for numeric variables.
+`r if (knitr::is_latex_output()) "@tbl-4-pdf shows a summary table for numeric variables." else ""`
+
+`r if (knitr::is_html_output()) "@tbl-4-html shows a summary table for numeric variables." else ""`
 
@@
                  histogram=T,
+                 output="flextable", # adjust to html for HTML output
                  title = "Summary: Numeric variables")
@@
 
-```{r eval=F}
-#| label: tbl-4
+```{r eval = knitr::is_latex_output()}
+#| label: tbl-4-pdf
 #| tbl-cap: "Summary: Numeric variables"
@@
                  histogram=T, 
-                 output="default",
+                 output="flextable",
+                 title = "Summary: Numeric variables")
+```
+
+```{r eval = knitr::is_html_output()}
+#| label: tbl-4-html
+#| tbl-cap: "Summary: Numeric variables"
+library(modelsummary)
+datasummary_skim(swiss, 
+                 type="numeric", 
+                 histogram=T, 
+                 output="html",
                  title = "Summary: Numeric variables")
@@
                             type="categorical", 
-                            output = 'flextable')
+                            output = 'flextable')  # adjust to html for HTML output
 
@@
 #| tbl-cap: "Summary: Categorical variables"
-#| results: "asis"
+
 # Create categorical variables
@@
 library(flextable)
-library(modelsummary)
 tab_cat <- datasummary_skim(swiss, 
@@
 
-# additionally we want to change the font, fontsize and spacing
+#additionally we want to change the font, fontsize and spacing
 library("gdtools")
@@
 
-@tbl-6 shows the output for a regression table. Make sure you name all your models and explicitly refer to model names (M1, M2 etc.) in the text.
+`r if (knitr::is_latex_output()) "@tbl-6-pdf shows a summary table for numeric variables." else ""`
 
-```{r echo=TRUE, paged.print=FALSE, results="asis"}
+`r if (knitr::is_html_output()) "@tbl-6-html shows a summary table for numeric variables." else ""`
+
+```{{r}}
 #| label: tbl-6
-#| tbl-cap: "Linear regression"
+#| tbl-cap: "Linear regression with stargazer"
+#| results: "asis"
 library(stargazer)
@@
 stargazer(M1, M2, M3,
-          title = "Regression table with stargazer",
-          label="tab2", 
+          type="latex", # adjust to html for HTML output
           table.placement = "H", 
@@
 
-## Regression table with modelsummary
+```{r eval = knitr::is_latex_output()}
+#| label: tbl-6-pdf
+#| tbl-cap: "Linear regression with stargazer"
+#| results: "asis"
+library(stargazer)
+M1 <- lm(Fertility ~ Education + Agriculture, data = swiss)
+M2 <- lm(Fertility ~ Education + Catholic, data = swiss)
+M3 <- lm(Fertility ~ Education + Infant.Mortality + Agriculture, data = swiss)
+stargazer(M1, M2, M3,
+          type="latex", # adjust to html for HTML output
+          table.placement = "H", 
+          column.labels = c("M1", "M2", "M3"),
+          model.numbers = FALSE,
+          header=FALSE)
+```
 
-@tbl-7 shows the output for a regression table. Make sure you name all the models you estimate (even if its 50) and explicitly refer to model names (M1, M2 etc.) in the text.
+```{r eval = knitr::is_html_output()}
+#| label: tbl-6-html
+#| tbl-cap: "Linear regression with stargazer"
+#| results: "asis"
+library(stargazer)
+M1 <- lm(Fertility ~ Education + Agriculture, data = swiss)
+M2 <- lm(Fertility ~ Education + Catholic, data = swiss)
+M3 <- lm(Fertility ~ Education + Infant.Mortality + Agriculture, data = swiss)
+stargazer(M1, M2, M3,
+          type="html",
+          table.placement = "H", 
+          column.labels = c("M1", "M2", "M3"),
+          model.numbers = FALSE,
+          header=FALSE)
+```
 
-```{r eval=F}
+`r if (knitr::is_html_output()) "## Regression table with modelsummary" else ""`
+
+`r if (knitr::is_html_output()) "@tbl-7 (HTML only) shows the output for a regression table.
+Make sure you name all the models you estimate (even if its 50) and explicitly refer to model names (M1, M2 etc.) in the text." else ""`
+
+```{{r}}
 #| label: tbl-7
-#| tbl-cap: "Linear regression"
+#| tbl-cap: "Linear regression with modelsummary"
+library(modelsummary)
+M1 <- lm(Fertility ~ Education + Agriculture, data = swiss)
+M2 <- lm(Fertility ~ Education + Catholic, data = swiss)
+M3 <- lm(Fertility ~ Education + Infant.Mortality + Agriculture, data = swiss)
+models <- list("M1" = M1, "M2" =  M2, "M3" = M3)
+
+library(gt)
+# additionally we want to change the font, font size and spacing
+modelsummary(models,
+             output = 'gt',
+             notes = "Notes: some notes...") %>%
+    tab_spanner(label = 'Dependent variable: Fertility', columns = 2:4) %>%
+  tab_options(
+    table.font.size = 10,
+    data_row.padding = px(1),
+    table.border.top.color = "white",
+    heading.border.bottom.color = "black",
+    row_group.border.top.color = "black",
+    row_group.border.bottom.color = "white",
+    table.border.bottom.color = "white",
+    column_labels.border.top.color = "black",
+    column_labels.border.bottom.color = "black",
+    table_body.border.bottom.color = "black",
+    table_body.hlines.color = "white"
+  )
+```
+
+```{r eval = knitr::is_html_output()}
+#| label: tbl-7
+#| tbl-cap: "Linear regression with modelsummary"
 library(modelsummary)

### 2023-03-11 paper.qmd
add mode source always

@@
   message: false
+editor:
+  mode: source
 ---
@@
 # Why reproducible research (in R)?
+
 Some arguments...
 
-* **Access**: Research is normally funded by taxpayers (researchers are also taxpayers).
+-   **Access**: Research is normally funded by taxpayers (researchers are also taxpayers).
 Hence, it should be freely accessible to everyone without any barriers, e.g., without requiring commercial software.
@@
 
-* **Reproducability**: Even if you have written a study and analyzed the data yourself you will forget what you did after a few months.
+-   **Reproducability**: Even if you have written a study and analyzed the data yourself you will forget what you did after a few months.
 A fully reproducible setup will help you to trace back your own steps.
@@
 
-* **Errors**: Manual steps in data analysis (e.g., manually copy/pasting values into a table etc.) may introduce errors. Quarto allows you to **automatize** such steps and/or avoid them.
+-   **Errors**: Manual steps in data analysis (e.g., manually copy/pasting values into a table etc.) may introduce errors. Quarto allows you to **automatize** such steps and/or avoid them.
 
-* **Revisions**: Revising a paper takes much less time if you have all the code you need in one place, i.e., one `.qmd` file.
+-   **Revisions**: Revising a paper takes much less time if you have all the code you need in one place, i.e., one `.qmd` file.
 For instance, if you decide to exclude a subset of your data you simply need to insert one line of your code at the beginning and everything is rebuilt/re-estimated automatically.
@@
 [Quarto](https://quarto.org/) is an open-source, scientific and technical publishing system build on `pandoc`.
-Their creators from RStudio also call it [„the next generation of R Markdown“](https://quarto.org/docs/computations/r.html).
+Their creators from RStudio also call it [„the next generation of R Markdown"](https://quarto.org/docs/computations/r.html).\
 Like RMarkdown, Quarto uses `Knitr` to execute `R` code, and is therefore able to render most existing `.rmd` files without having to make any adjustments.
-As in `rmd` and other more recent developments like R's `pagedown`-package, one of the main goals of Quarto is to weave together narrative text and code and produce nicely formatted output documents.
+As in `rmd` and other more recent developments like R's `pagedown`-package, one of the main goals of Quarto is to weave together narrative text and code and produce nicely formatted output documents.\
 Differences to RMarkdown are not only but especially due to two facts: first, Quarto is multi-lingual (R, Python, Julia, Observable JS, etc.) and second in addition to RMarkdown's workflow where the final conversion layer uses `pandoc` to create documents, Quarto uses pandoc with Lua filters (which include latex in pandoc metadata).
@@
 Multi-language means you can use R, Python, Julia and other languages to create quarto documents.
-In this template we’ll focus on showing you how to use **RStudio with Quarto**.
+In this template we'll focus on showing you how to use **RStudio with Quarto**.
 
@@
 
-* ...install [R](https://www.r-project.org/) and [Rstudio](https://posit.co/download/rstudio-desktop/) (important: most recent versions, e.g. RStudio v2022.07) [@R2017; @Rstudio2015].
-* ...install the "rmarkdown"-package from CRAN using the code below [@markdown2017].
+-   ...install [R](https://www.r-project.org/) and [Rstudio](https://posit.co/download/rstudio-desktop/) (important: most recent versions, e.g. RStudio v2022.07) [@R2017; @Rstudio2015].
+-   ...install the "rmarkdown"-package from CRAN using the code below [@markdown2017].
 
-```{{r}}
+``` {{r}}
 install.packages("rmarkdown")
@@
 ```
-* ...install [tinytex](https://yihui.name/tinytex/), a lightweight version of Tex Live [@markdown2017; @tinytex] in order to compile PDFs.
 
-```{{r}}
+-   ...install [tinytex](https://yihui.name/tinytex/), a lightweight version of Tex Live [@markdown2017; @tinytex] in order to compile PDFs.
+
+``` {{r}}
 install.packages('tinytex')
@@
 
-* ...also install the packages below using the code below [@knitr1; @knitr2; @knitr3; @kableextra; @hlavac2013stargazer; @modelsummary; @gt].
+-   ...also install the packages below using the code below [@knitr1; @knitr2; @knitr3; @kableextra; @hlavac2013stargazer; @modelsummary; @gt].
 
-```{{r}}
+``` {{r}}
 install.packages(c("knitr", "kableExtra",
@@
 ```
-* ...download the 3 input files we created --- `paper.qmd`, `references.bib` and `data.csv` --- from [this folder](https://drive.google.com/drive/u/0/folders/1eT_9WaineZRB-nw-cQhNOy16sWhd_t2s).
+
+-   ...download the 3 input files we created --- `paper.qmd`, `references.bib` and `data.csv` --- from [this folder](https://drive.google.com/drive/u/0/folders/1eT_9WaineZRB-nw-cQhNOy16sWhd_t2s).
 Ignore the other files.
-* ...finally store all files from above together in one folder (and use this folder as your working directory later on).
+-   ...finally store all files from above together in one folder (and use this folder as your working directory later on).
 
@@
 
-* `paper.qmd` (the underlying R Markdown file).
-* `references.bib` (the bibliography).
-    + We use paperpile to manage references and export the `.bib` file into the folder that contains the `.qmd` file.
-* `data.csv` (some raw data).
+-   `paper.qmd` (the underlying R Markdown file).
+-   `references.bib` (the bibliography).
+    -   We use paperpile to manage references and export the `.bib` file into the folder that contains the `.qmd` file.
+-   `data.csv` (some raw data).
 
@@
 Close R/Rstudio and directly open `paper.qmd` with RStudio.
-Doing so assures that the working directory is set to the folder that contains `paper.qmd` and the other files.^[You can always check your working directory in R with `getwd()`.]
+Doing so assures that the working directory is set to the folder that contains `paper.qmd` and the other files.[^1]
+
+[^1]: You can always check your working directory in R with `getwd()`.
 
@@
 If you want your document to be rendered in multiple output formats (e.g., HTML and PDF like we did here), you need to specify both outputs in your YAML header.
-Take `paper.qmd` (the underlying quarto file of this pdf) and have a look at the YAML (line \#14 - \#16) to see how to specify different outputs.
+Take `paper.qmd` (the underlying quarto file of this pdf) and have a look at the YAML (line #14 - #16) to see how to specify different outputs.
 
@@
 `paper.qmd` (the underlying quarto file of this PDF) will show you how we referenced @sec-tables right here namely with '`@sec-tables`' whereas the corresponding section title was assigned the corresponding label '`# Tables {#sec-tables}`'.
-Pay attention that when using section cross-references, you need to enable the number-sections option in your YAML (line \#22 of this `qmd` file).
+Pay attention that when using section cross-references, you need to enable the number-sections option in your YAML (line #22 of this `qmd` file).
 
@@
 
-```{{r}}
+``` {{r}}
 #|label: fig-versioning
@@
 # Data
+
 ## Import
@@
 
-```{{r}}
+``` {{r}}
 x <- 1:10
@@
 
-```{{r}}
+``` {{r}}
 |echo=T
@@
 ## Putting your entire data into the .qmd file
+
 Applying the function `dput()` to an object gives you the code needed to reproduce that object.
@@
 
-```{{r}}
+``` {{r}}
 dput(data[1:5,]) # here we only take a subset
@@
 
-```{{r}}
+``` {{r}}
 data <- structure(list(Fertility = c(80.2, 83.1, 92.5, 85.8, 76.9), Agriculture = c(17, 
@@
 
-```{{r}}
+``` {{r}}
 #| label: tbl-1
@@
 
-```{{r}}
+``` {{r}}
 #| label: tbl-2
@@
 
-```{{r}}
+``` {{r}}
 #| label: tbl-4
@@
 
-```{{r}}
+``` {{r}}
 #| label: tbl-5
@@
 
-```{{r}}
+``` {{r}}
 #| label: tbl-6
@@
 
-```{{r}}
+``` {{r}}
 #| label: tbl-7
@@
 `r rinline("round(coef(M1)[2], 2)")` will insert the coefficient as follows: `r round(coef(M1)[2], 2)`.
-Or `r rinline("3 + 7")` will insert a `r 3 + 7` in the text.
+Or `r rinline("3 + 7")` will insert a `r 3 + 7` in the text.\
 Inline code/results that depend on earlier objects in your document will automatically be updated once you change those objects.
@@
 You can simply do so in the beginning of your code and push play subsequently..
-at time you might have to set `cache = FALSE` at the beginning so that all the code chunks are rerun.
+at time you might have to set `cache = FALSE` at the beginning so that all the code chunks are rerun.\
 Researchers often avoid referring to results in-text etc.
@@
 ## R base graphs
+
 A fully reproducible manuscript would create diagrams directly in the `.qmd` file and insert them in the appropriate place. It's relatively simple to insert R base graphs as you can see in @fig-1.
 
-```{{r}}
+``` {{r}}
 #| label: fig-1
@@
 
-```{{r}}
+``` {{r}}
 #| label: fig-2
@@
 
-* first, use the Render Button in R Studio.
+-   first, use the Render Button in R Studio.
 If you click it without using the drop down menu, by default it will create HTML since it is the first format listed in our YAML.
@@
 
-* second, if you are not using RStudio and/or you prefer to render from the R console, use the `quarto` package to render to all formats at the same time with the `quarto_render()` function:
+-   second, if you are not using RStudio and/or you prefer to render from the R console, use the `quarto` package to render to all formats at the same time with the `quarto_render()` function:
 
-```{{r}}
+``` {{r}}
 quarto::quarto_render()
@@
 
-```{{r}}
+``` {{r}}
 install.packages("quarto")
@@
 
-```{{r}}
+``` {{r}}
 quarto::quarto_render(
@@
 # Good practices for reproducibility
+
 Every researcher has his own optimized setup. Currently we would recommend the following:
 
-* Keep all files of your project (that matter for producing the PDF) in one folder without subfolders.
-You can zip and directly upload that folder to the [Harvard dataverse](https://dataverse.harvard.edu/).^[Another good folder setup would be to store all files needed as input files for the R Markdown manuscript in a subfolder called "input" and all output files that are produced apart from paper.html and paper.pdf in a subfolder called "output".]
-* Make sure that filenames have a logic to them.
-    + Main file with text/code: "paper.qmd", "report.qmd" 
-    + Data files: "data_xxxxxx.*"
-    + Image files: "fig_xxxxxx.*"
-    + Tables files: "table_xxxx.*"
-    + etc.
-    + Ideally, your filenames will correspond to the names in the paper. For instance, Figure 1 in the paper may have a corresponding file called `fig_1_xxxxx.pdf`.
-* Use the document outline in R studio (Ctrl + Shift + O) when you work with R Markdown.
-* Name rchunks according to what they do or produce:
-    + "`fig-...`" for chunks producing figures
-    + "`table-...`" for chunks producing tables
-    + "`model-...`" for chunks producing model estimates
-    + "`import-...`" for chunks importing data
-    + "`recoding-...`" for chunks in which data is recoded
-* Use "really" informative variable names:
-    + Q: What do you think does the variable *trstep* measure? It actually measures trust in the European parliament. 
-        + How could we call this variable instead? Yes, `trust.european.parliament` which is longer but will probably be understood by another researcher in 50 years. 
-    + If your setup is truly reproducible you will probably re-use the variable names that you generate as variable names in the tables you produce. Hence, there is an incentive to use good names.
-* Use unique identifiers in the final R Markdown document paper.qmd that you upload:
-    + Think of someone who wants to produce Figure 1/Model 1/Table 1 in your paper but doesn't find it in your code...
-        + Name the chunks "fig-1", "fig-2" as the are named in the published paper.
-        + Name the chunks that produce tables "table-1", "table-2" etc. as they are named in the published paper.
-        + Name your statistical models in your R code "M1", "M2" as they are named in the published paper.
+-   Keep all files of your project (that matter for producing the PDF) in one folder without subfolders.
+You can zip and directly upload that folder to the [Harvard dataverse](https://dataverse.harvard.edu/).[^2]
+-   Make sure that filenames have a logic to them.
+    -   Main file with text/code: "paper.qmd", "report.qmd"
+    -   Data files: "data_xxxxxx.\*"
+    -   Image files: "fig_xxxxxx.\*"
+    -   Tables files: "table_xxxx.\*"
+    -   etc.
+    -   Ideally, your filenames will correspond to the names in the paper. For instance, Figure 1 in the paper may have a corresponding file called `fig_1_xxxxx.pdf`.
+-   Use the document outline in R studio (Ctrl + Shift + O) when you work with R Markdown.
+-   Name rchunks according to what they do or produce:
+    -   "`fig-...`" for chunks producing figures
+    -   "`table-...`" for chunks producing tables
+    -   "`model-...`" for chunks producing model estimates
+    -   "`import-...`" for chunks importing data
+    -   "`recoding-...`" for chunks in which data is recoded
+-   Use "really" informative variable names:
+    -   Q: What do you think does the variable *trstep* measure? It actually measures trust in the European parliament.
+        -   How could we call this variable instead? Yes, `trust.european.parliament` which is longer but will probably be understood by another researcher in 50 years.
+    -   If your setup is truly reproducible you will probably re-use the variable names that you generate as variable names in the tables you produce. Hence, there is an incentive to use good names.
+-   Use unique identifiers in the final R Markdown document paper.qmd that you upload:
+    -   Think of someone who wants to produce Figure 1/Model 1/Table 1 in your paper but doesn't find it in your code...
+        -   Name the chunks "fig-1", "fig-2" as the are named in the published paper.
+        -   Name the chunks that produce tables "table-1", "table-2" etc. as they are named in the published paper.
+        -   Name your statistical models in your R code "M1", "M2" as they are named in the published paper.
+
+[^2]: Another good folder setup would be to store all files needed as input files for the R Markdown manuscript in a subfolder called "input" and all output files that are produced apart from paper.html and paper.pdf in a subfolder called "output".
 
 # Additional tricks for publishing
-* Make your script anonymous
-    + Simply put a `<!-- ... -->` around any identifying information, e.g., author names, title footnote etc.
-* Counting words
-    + Use adobe acrobat (commerical software) to convert your file to a word file.
+
+-   Make your script anonymous
+    -   Simply put a `<!-- ... -->` around any identifying information, e.g., author names, title footnote etc.
+-   Counting words
+    -   Use adobe acrobat (commerical software) to convert your file to a word file.
 Then open in word and delete all the parts that shouldn't go into the word count.
 The word count is displayed in the lower right.
-    + Use an one of the online services to count your words (search for "pdf word count")
-* Appendix: You can change the numbering format for the appendix in the rmd file
-    + What is still not possible in this document is to automatically have separate reference sections for paper and appendix.
-* Journals may require you to use their tex style: Sometimes you can simply use their template in your rmarkdown file.
+    -   Use an one of the online services to count your words (search for "pdf word count")
+-   Appendix: You can change the numbering format for the appendix in the rmd file
+    -   What is still not possible in this document is to automatically have separate reference sections for paper and appendix.
+-   Journals may require you to use their tex style: Sometimes you can simply use their template in your rmarkdown file.
 See [here](https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/LDUMNY) for a PLOS one example.
@@
 # Citation styles
+
 If your study needs to follow a particular citation style, you can set the corresponding style in the header of your `.qmd` document. To do so you have to download the corresponding `.csl` file.
@@
 
-# References {-}
+# References {.unnumbered}
 
 ::: {#refs}
-
 :::
@@
 Then  copy the output to a .bib file and add a unique citation key -->
+
 <!-- https://bookdown.org/yihui/rmarkdown-cookbook/write-bib.html -->
@@
 
-# Online appendix {-}
+# Online appendix {.unnumbered}
 

### 2023-03-11 paper.qmd
Add cache = false

@@
   message: false
+  cache: false
 editor:

### 2023-03-20 paper.qmd
Update files after successful compilation
Update tables for both outputs.

@@
     keeptex: true
+    include-in-header: 
+      text: |
+        \usepackage{booktabs}
+        \usepackage{siunitx}
+        \newcolumntype{d}{S[
+            input-open-uncertainty=,
+            input-close-uncertainty=,
+            parse-numbers = false,
+            table-align-text-pre=false,
+            table-align-text-post=false
+         ]}
 date: 'last-modified'
@@
 install.packages(c("knitr", "kableExtra",
-                   "stargazer", "modelsummary", "gt"))
+                   "stargazer", "modelsummary", "gt", "gtsummary"))
 ```
@@
 install.packages(c("knitr", "kableExtra",
-                   "stargazer", "modelsummary", "knitr", "gt"))
+                   "stargazer", "modelsummary", "knitr", "gt", "gtsummary"))
 ```
@@
 
-## Tables with stargazer()
-
-`r if (knitr::is_latex_output()) "@tbl-2-pdf was generated using the stargazer() R package [@hlavac2013stargazer] and shows summary statistics of our data." else ""`
-
-`r if (knitr::is_html_output()) "@tbl-2-html was generated using the stargazer() R package [@hlavac2013stargazer] and shows summary statistics of our data." else ""`
-
-``` {{r}}
-#| label: tbl-2
-#| tbl-cap: "Summary table with stargazer"
-#| results: "asis"
-library(stargazer)
-stargazer(swiss, 
-          table.placement = "H", 
-          type="latex", # adjust to html for HTML output
-          header=FALSE)
-```
-
-```{r, eval = knitr::is_latex_output()}
-#| label: tbl-2-pdf
-#| tbl-cap: "Summary table with stargazer"
-#| results: "asis"
-library(stargazer)
-stargazer(swiss, 
-          table.placement = "H", 
-          type="latex",
-          header=FALSE)
-```
-
-```{r eval = knitr::is_html_output()}
-#| label: tbl-2-html
-#| tbl-cap: "Summary table with stargazer"
-#| results: "asis"
-library(stargazer)
-stargazer(swiss,
-          table.placement = "H", 
-          type="html",
-          header=FALSE)
-```
-
-In the above example, note that the `type` argument defaults to `latex` and you have to insert `html` if you desire HTML output.
-Lastly, note, how we specified 'results: asis' into the code chunk options.
-This is important so that latex is properly printed to the PDF.
-
 ## Tables with modelsummary
@@
 
-`r if (knitr::is_latex_output()) "@tbl-4-pdf shows a summary table for numeric variables." else ""`
-
-`r if (knitr::is_html_output()) "@tbl-4-html shows a summary table for numeric variables." else ""`
+@tbl-4 shows a summary table for numeric variables.
 
@@
 
-```{r eval = knitr::is_latex_output()}
-#| label: tbl-4-pdf
+```{r}
+#| label: tbl-4
 #| tbl-cap: "Summary: Numeric variables"
+
+if(knitr::is_latex_output()){outputformat <- "flextable"}
+if(knitr::is_html_output()){outputformat <- "html"}
+
 library(modelsummary)
@@
                  histogram=T, 
-                 output="flextable",
-                 title = "Summary: Numeric variables")
-```
-
-```{r eval = knitr::is_html_output()}
-#| label: tbl-4-html
-#| tbl-cap: "Summary: Numeric variables"
-library(modelsummary)
-datasummary_skim(swiss, 
-                 type="numeric", 
-                 histogram=T, 
-                 output="html",
+                 output= outputformat,
                  title = "Summary: Numeric variables")
@@
 
-## Regression Table with stargazer
+### Regression table with modelsummary
 
-`r if (knitr::is_latex_output()) "@tbl-6-pdf shows a summary table for numeric variables." else ""`
-
-`r if (knitr::is_html_output()) "@tbl-6-html shows a summary table for numeric variables." else ""`
-
-``` {{r}}
-#| label: tbl-6
-#| tbl-cap: "Linear regression with stargazer"
-#| results: "asis"
-library(stargazer)
-M1 <- lm(Fertility ~ Education + Agriculture, data = swiss)
-M2 <- lm(Fertility ~ Education + Catholic, data = swiss)
-M3 <- lm(Fertility ~ Education + Infant.Mortality + Agriculture, data = swiss)
-stargazer(M1, M2, M3,
-          type="latex", # adjust to html for HTML output
-          table.placement = "H", 
-          column.labels = c("M1", "M2", "M3"),
-          model.numbers = FALSE,
-          header=FALSE)
-```
-
-```{r eval = knitr::is_latex_output()}
-#| label: tbl-6-pdf
-#| tbl-cap: "Linear regression with stargazer"
-#| results: "asis"
-library(stargazer)
-M1 <- lm(Fertility ~ Education + Agriculture, data = swiss)
-M2 <- lm(Fertility ~ Education + Catholic, data = swiss)
-M3 <- lm(Fertility ~ Education + Infant.Mortality + Agriculture, data = swiss)
-stargazer(M1, M2, M3,
-          type="latex", # adjust to html for HTML output
-          table.placement = "H", 
-          column.labels = c("M1", "M2", "M3"),
-          model.numbers = FALSE,
-          header=FALSE)
-```
-
-```{r eval = knitr::is_html_output()}
-#| label: tbl-6-html
-#| tbl-cap: "Linear regression with stargazer"
-#| results: "asis"
-library(stargazer)
-M1 <- lm(Fertility ~ Education + Agriculture, data = swiss)
-M2 <- lm(Fertility ~ Education + Catholic, data = swiss)
-M3 <- lm(Fertility ~ Education + Infant.Mortality + Agriculture, data = swiss)
-stargazer(M1, M2, M3,
-          type="html",
-          table.placement = "H", 
-          column.labels = c("M1", "M2", "M3"),
-          model.numbers = FALSE,
-          header=FALSE)
-```
-
-`r if (knitr::is_html_output()) "## Regression table with modelsummary" else ""`
-
-`r if (knitr::is_html_output()) "@tbl-7 (HTML only) shows the output for a regression table.
-Make sure you name all the models you estimate (even if its 50) and explicitly refer to model names (M1, M2 etc.) in the text." else ""`
+@tbl-7 shows the output for a regression table.
+Make sure you name all the models you estimate with unique identifiers (even if its 50 models) and explicitly refer to model names (M1, M2 etc.) in the text.
 
@@
 library(gt)
+library(gtsummary)
 # additionally we want to change the font, font size and spacing
@@
 
-```{r eval = knitr::is_html_output()}
+```{r}
 #| label: tbl-7
 #| tbl-cap: "Linear regression with modelsummary"
+#| results: "asis"
+
 library(modelsummary)
@@
 
+# Hier Weiter
+
 library(gt)
+library(gtsummary)
 # additionally we want to change the font, font size and spacing
-modelsummary(models,
+table7 <- modelsummary(models,
              output = 'gt',
-             notes = "Notes: some notes...") %>%
-    tab_spanner(label = 'Dependent variable: Fertility', columns = 2:4) %>%
-  tab_options(
-    table.font.size = 10,
-    data_row.padding = px(1),
-    table.border.top.color = "white",
-    heading.border.bottom.color = "black",
-    row_group.border.top.color = "black",
-    row_group.border.bottom.color = "white",
-    table.border.bottom.color = "white",
-    column_labels.border.top.color = "black",
-    column_labels.border.bottom.color = "black",
-    table_body.border.bottom.color = "black",
-    table_body.hlines.color = "white"
-  )
+             stars = TRUE,
+             notes = "Notes: put some notes here...",
+             threeparttable = TRUE) %>%
+    tab_spanner(label = 'Dependent variable: Fertility', columns = 2:4)
+
+if(knitr::is_html_output()){table7}else{table7 %>% gt::as_latex() %>% as.character() %>% gsub("\\(Intercept\\)", "Intercept", .) %>% cat()}
+
 ```

### 2023-03-20 paper.qmd
Small edits.

@@
                  histogram=T,
-                 output="flextable", # adjust to html for HTML output
+                 output="flextable",
                  title = "Summary: Numeric variables")
@@
 
-if(knitr::is_latex_output()){outputformat <- "flextable"}
-if(knitr::is_html_output()){outputformat <- "html"}
-
 library(modelsummary)
@@
                  histogram=T, 
-                 output= outputformat,
+                 output= "flextable",
                  title = "Summary: Numeric variables")
@@
                             type="categorical", 
-                            output = 'flextable')  # adjust to html for HTML output
+                            output = 'flextable')
 
@@
   fontsize(size=12, part="header") %>%
-  fontsize(size=10, j=1:4) %>%
+  fontsize(size=12, j=1:4) %>%
   line_spacing(space = 0.3, part = "all")
@@
   )
+
+# Produce right output depending on format (html vs. latex)
+if(knitr::is_html_output()){table7}else{table7 %>% gt::as_latex() %>% as.character() %>% gsub("\\(Intercept\\)", "Intercept", .) %>% cat()}
 ```
@@
 if(knitr::is_html_output()){table7}else{table7 %>% gt::as_latex() %>% as.character() %>% gsub("\\(Intercept\\)", "Intercept", .) %>% cat()}
-
 ```

### 2023-03-23 paper.qmd
remove all references to stargazer.

@@
 
--   ...also install the packages below using the code below [@knitr1; @knitr2; @knitr3; @kableextra; @hlavac2013stargazer; @modelsummary; @gt].
+-   ...also install the packages below using the code below [@knitr1; @knitr2; @knitr3; @kableextra; @modelsummary; @gt].
 
 ``` {{r}}
-install.packages(c("knitr", "kableExtra",
-                   "stargazer", "modelsummary", "gt", "gtsummary"))
+install.packages(c("knitr", "kableExtra", "modelsummary", "gt", "gtsummary"))
 ```
@@
 install.packages(c("knitr", "kableExtra",
-                   "stargazer", "modelsummary", "knitr", "gt", "gtsummary"))
+                   "modelsummary", "knitr", "gt", "gtsummary"))
 ```
@@
 ```
-
-<!-- the stargazer section is only shown when the output format is PDF and not HTML -->
 

### 2023-03-25 paper.qmd
remove stargazer package + resolve overlap in latex modelsummary tables.

@@
 To further deepen your understanding of the different components underlying this template, go on and learn about R, [Markdown](https://en.wikipedia.org/wiki/Markdown), [R Markdown](https://rmarkdown.rstudio.com/lesson-1.html) and [Quarto](https://quarto.org/).
-For Quarto also consider reading their [FAQ](https://quarto.org/docs/faq/rmarkdown.html).
+For more information on Quarto also consider reading their [FAQ](https://quarto.org/docs/faq/rmarkdown.html).
 
@@
 
-@tbl-4 shows a summary table for numeric variables.
+@tbl-2 shows a summary table for numeric variables.
 
 ``` {{r}}
-#| label: tbl-4
+#| label: tbl-2
 #| tbl-cap: "Summary: Numeric variables"
@@
                  type="numeric", 
-                 histogram=T,
-                 output="flextable",
-                 title = "Summary: Numeric variables")
+                 histogram=F,
+                 output="flextable") %>% 
+  # column widths
+  flextable::autofit()
 ```
@@
 ```{r}
-#| label: tbl-4
+#| label: tbl-2
 #| tbl-cap: "Summary: Numeric variables"
@@
                  type="numeric", 
-                 histogram=T, 
-                 output= "flextable",
-                 title = "Summary: Numeric variables")
+                 histogram=F, 
+                 output= "flextable") %>% 
+  # column widths
+  flextable::width(1, width = 1) %>% 
+  flextable::width(2:8, width = 0.5)
+
 ```
@@
 
-@tbl-5 shows a summary table for categorical variables.
+@tbl-3 shows a summary table for categorical variables. Notice how for this table we also changed the font to be "Times New Roman".
 
 ``` {{r}}
-#| label: tbl-5
+#| label: tbl-3
 #| tbl-cap: "Summary: Categorical variables"
@@
   fontsize(size=10, j=1:4) %>% 
-  line_spacing(space = 0.3, part = "all")
+  line_spacing(space = 0.3, part = "all") %>% 
+  # column widths
+  flextable::autofit()
 
@@
 ```{r}
-#| label: tbl-5
+#| label: tbl-3
 #| tbl-cap: "Summary: Categorical variables"
@@
   fontsize(size=12, j=1:4) %>%
-  line_spacing(space = 0.3, part = "all")
+  line_spacing(space = 0.3, part = "all") %>% 
+  # column widths
+  flextable::autofit()
 
@@
 
-@tbl-7 shows the output for a regression table.
+@tbl-4 shows the output for a regression table.
 Make sure you name all the models you estimate with unique identifiers (even if its 50 models) and explicitly refer to model names (M1, M2 etc.) in the text.
@@
 ``` {{r}}
-#| label: tbl-7
+#| label: tbl-4
 #| tbl-cap: "Linear regression with modelsummary"
@@
 # Produce right output depending on format (html vs. latex)
-if(knitr::is_html_output()){table7}else{table7 %>% gt::as_latex() %>% as.character() %>% gsub("\\(Intercept\\)", "Intercept", .) %>% cat()}
+if(knitr::is_html_output()){table4}else{table4 %>% gt::as_latex() %>% as.character() %>% gsub("\\(Intercept\\)", "Intercept", .) %>% cat()}
 ```
@@
 ```{r}
-#| label: tbl-7
+#| label: tbl-4
 #| tbl-cap: "Linear regression with modelsummary"
 #| results: "asis"
-
 library(modelsummary)
@@
 
-# Hier Weiter
-
 library(gt)
@@
 # additionally we want to change the font, font size and spacing
-table7 <- modelsummary(models,
+table4 <- modelsummary(models,
              output = 'gt',
@@
 
-if(knitr::is_html_output()){table7}else{table7 %>% gt::as_latex() %>% as.character() %>% gsub("\\(Intercept\\)", "Intercept", .) %>% cat()}
+if(knitr::is_html_output()){table4}else{table4 %>% gt::as_latex() %>% as.character() %>% gsub("\\(Intercept\\)", "Intercept", .) %>% cat()}
 ```

### 2023-03-30 paper.qmd
Small updates (typos, date)

@@
 date: 'last-modified'
-date-format: '[First version: February  16, 2023 <br>This version:] MMMM D, YYYY'
+date-format: '[This version:] MMMM D, YYYY [<br>(First version: February  16, 2023)]'
 abstract: 'This paper provides a template to write a reproducible scientific paper with RStudio and Quarto.
-It falls into a series on other templates, namely on RMarkdown and pagedown that can be downloaded [here](https://osf.io/k8jhx/) [@Bauer2018-dl] and [here](https://osf.io/q395s) [@Bauer2021-dl].
+It falls into a series of other templates, namely using RMarkdown [@Bauer2018-dl] (see [here](https://osf.io/k8jhx/)) and Pagedown [@Bauer2021-dl] (see [here](https://osf.io/q395s)).
 Below we outline some of the "tricks"/code (e.g., referencing tables, sections etc.) we had to figure out to produce this document.
-The underlying files which produce this document can be [downloaded here](https://drive.google.com/drive/u/0/folders/1eT_9WaineZRB-nw-cQhNOy16sWhd_t2s).
+The underlying files which produce this document can be [downloaded here](https://drive.google.com/drive/folders/1jDPpXdX6dusXpKkXEhDGujWDGf52TaxD?usp=sharing).
 We are convinced that in the future there will be many improvements and developments with regards to RStudio and Quarto.
@@
 ``` {{r}}
-data <- structure(list(Fertility = c(80.2, 83.1, 92.5, 85.8, 76.9), Agriculture = c(17, 
-45.1, 39.7, 36.5, 43.5), Examination = c(15L, 6L, 5L, 12L, 17L
-), Education = c(12L, 9L, 5L, 7L, 15L), Catholic = c(9.96, 84.84, 
-93.4, 33.77, 5.16), Infant.Mortality = c(22.2, 22.2, 20.2, 20.3, 
-20.6)), class = "data.frame", row.names = c(NA, -5L))
+data <- structure(list(Fertility = c(80.2, 83.1, 92.5, 85.8, 76.9), 
+                       Agriculture = c(17, 45.1, 39.7, 36.5, 43.5), 
+                       Examination = c(15L, 6L, 5L, 12L, 17L), 
+                       Education = c(12L, 9L, 5L, 7L, 15L), 
+                       Catholic = c(9.96, 84.84, 93.4, 33.77, 5.16), 
+                       Infant.Mortality = c(22.2, 22.2, 20.2, 20.3, 20.6)), 
+                       class = "data.frame", row.names = c(NA, -5L))
 ```
@@
 ```{r}
-data <- structure(list(Fertility = c(80.2, 83.1, 92.5, 85.8, 76.9), Agriculture = c(17, 
-45.1, 39.7, 36.5, 43.5), Examination = c(15L, 6L, 5L, 12L, 17L
-), Education = c(12L, 9L, 5L, 7L, 15L), Catholic = c(9.96, 84.84, 
-93.4, 33.77, 5.16), Infant.Mortality = c(22.2, 22.2, 20.2, 20.3, 
-20.6)), class = "data.frame", row.names = c(NA, -5L))
+data <- structure(list(Fertility = c(80.2, 83.1, 92.5, 85.8, 76.9),
+                       Agriculture = c(17, 45.1, 39.7, 36.5, 43.5), 
+                       Examination = c(15L, 6L, 5L, 12L, 17L), 
+                       Education = c(12L, 9L, 5L, 7L, 15L), 
+                       Catholic = c(9.96, 84.84, 93.4, 33.77, 5.16), 
+                       Infant.Mortality = c(22.2, 22.2, 20.2, 20.3, 20.6)), 
+                  class = "data.frame", row.names = c(NA, -5L))
 ```
@@
 # Produce right output depending on format (html vs. latex)
-if(knitr::is_html_output()){table4}else{table4 %>% gt::as_latex() %>% as.character() %>% gsub("\\(Intercept\\)", "Intercept", .) %>% cat()}
+if(knitr::is_html_output()){table4}else
+    {table4 %>% gt::as_latex() %>% 
+    as.character() %>% 
+    gsub("\\(Intercept\\)", "Intercept", .) %>% 
+    cat()}
 ```
@@
 
-if(knitr::is_html_output()){table4}else{table4 %>% gt::as_latex() %>% as.character() %>% gsub("\\(Intercept\\)", "Intercept", .) %>% cat()}
+if(knitr::is_html_output()){table4}else
+    {table4 %>% 
+    gt::as_latex() %>% 
+    as.character() %>% 
+    gsub("\\(Intercept\\)", "Intercept", .) %>% 
+    cat()}
 ```

### 2023-04-03 paper.qmd
Update links to gdrive.

@@
 
--   ...download the 3 input files we created --- `paper.qmd`, `references.bib` and `data.csv` --- from [this folder](https://drive.google.com/drive/u/0/folders/1eT_9WaineZRB-nw-cQhNOy16sWhd_t2s).
+-   ...download the 3 input files we created --- `paper.qmd`, `references.bib` and `data.csv` --- from [this folder](https://drive.google.com/drive/folders/1jDPpXdX6dusXpKkXEhDGujWDGf52TaxD?usp=sharing).
 Ignore the other files.
@@
 
-[Download these files](https://drive.google.com/drive/u/0/folders/1eT_9WaineZRB-nw-cQhNOy16sWhd_t2s) and save them into a folder.
+[Download these files](https://drive.google.com/drive/folders/1jDPpXdX6dusXpKkXEhDGujWDGf52TaxD?usp=sharing) and save them into a folder.
 Close R/Rstudio and directly open `paper.qmd` with RStudio.
@@
 
-Below we import an exemplary dataset (you can find `data.csv` in the [folder](https://drive.google.com/drive/u/0/folders/1eT_9WaineZRB-nw-cQhNOy16sWhd_t2s) with the other files).
+Below we import an exemplary dataset (you can find `data.csv` in the [folder](https://drive.google.com/drive/folders/1jDPpXdX6dusXpKkXEhDGujWDGf52TaxD?usp=sharing) with the other files).
 

### 2023-04-03 paper.qmd
Update files/links.

--- 2023-03-30 paper.qmd
@@
 Below we outline some of the "tricks"/code (e.g., referencing tables, sections etc.) we had to figure out to produce this document.
-The underlying files which produce this document can be [downloaded here](https://drive.google.com/drive/folders/1jDPpXdX6dusXpKkXEhDGujWDGf52TaxD?usp=sharing).
+The underlying files which produce this document can be [downloaded here](https://github.com/paulcbauer/Writing_a_reproducable_paper_with_quarto/) (click on Code -> Download ZIP).
 We are convinced that in the future there will be many improvements and developments with regards to RStudio and Quarto.
@@
 
--   ...download the 3 input files we created --- `paper.qmd`, `references.bib` and `data.csv` --- from [this folder](https://drive.google.com/drive/u/0/folders/1eT_9WaineZRB-nw-cQhNOy16sWhd_t2s).
-Ignore the other files.
+- ...download the 4 input files we created (see below) from [github](https://github.com/paulcbauer/Writing_a_reproducable_paper_with_quarto/) (click on Code -> Download ZIP).
+Ignore/delete the other files.
+
+		- `paper.qmd`
+		- `references.bib`
+		- `american-sociological-association.csl`
+		- `data.csv`
+
 - ...finally store all files from above together in one folder (and use this folder as your working directory later on).
@@
     -   We use paperpile to manage references and export the `.bib` file into the folder that contains the `.qmd` file.
+- `american-sociological-association.csl` (the bibliography style)
 -   `data.csv` (some raw data).
 
-[Download these files](https://drive.google.com/drive/u/0/folders/1eT_9WaineZRB-nw-cQhNOy16sWhd_t2s) and save them into a folder.
+[Download these files from github]((https://github.com/paulcbauer/Writing_a_reproducable_paper_with_quarto/) (click on Code -> Download ZIP) and save them into a folder.
 Close R/Rstudio and directly open `paper.qmd` with RStudio.
@@
 
-Below we import an exemplary dataset (you can find `data.csv` in the [folder](https://drive.google.com/drive/u/0/folders/1eT_9WaineZRB-nw-cQhNOy16sWhd_t2s) with the other files).
+Below we import an exemplary dataset (you can find `data.csv` in the [github repository]((https://github.com/paulcbauer/Writing_a_reproducable_paper_with_quarto/) with the other files).
 

### 2023-04-03 paper.qmd
Update with image for github code.

@@
 Below we outline some of the "tricks"/code (e.g., referencing tables, sections etc.) we had to figure out to produce this document.
-The underlying files which produce this document can be [downloaded here](https://github.com/paulcbauer/Writing_a_reproducable_paper_with_quarto/) (click on Code -> Download ZIP).
+The underlying files which produce this document can be [downloaded here](https://github.com/paulcbauer/Writing_a_reproducable_paper_with_quarto/) (click on ![](githubcode.png) $\Rightarrow$ **Download ZIP**).
 We are convinced that in the future there will be many improvements and developments with regards to RStudio and Quarto.
@@
 
-- ...download the 4 input files we created (see below) from [github](https://github.com/paulcbauer/Writing_a_reproducable_paper_with_quarto/) (click on Code -> Download ZIP).
+- ...download the 4 input files we created (see below) from [github](https://github.com/paulcbauer/Writing_a_reproducable_paper_with_quarto/) (click on ![](githubcode.png) $\Rightarrow$ **Download ZIP**).
 Ignore/delete the other files.
@@
 
-[Download these files from github]((https://github.com/paulcbauer/Writing_a_reproducable_paper_with_quarto/) (click on Code -> Download ZIP) and save them into a folder.
+[Download these files from github](https://github.com/paulcbauer/Writing_a_reproducable_paper_with_quarto/) (click on ![](githubcode.png) $\Rightarrow$ **Download ZIP**) and save them into a folder.
 Close R/Rstudio and directly open `paper.qmd` with RStudio.

### 2023-04-06 paper.qmd
Add results: "asis" to visible code of regression table.

@@
 #| tbl-cap: "Linear regression with modelsummary"
+#| results: "asis"
+
 library(modelsummary)

### 2023-04-07 paper.qmd
Adapt Section 3 to facilitate installation, drop html output for now because of path error

@@
 format:
-  html: default
   pdf:
@@
 We intend to update this file when we discover more convenient code.
-You can follow any updates on the [github repository](https://github.com/paulcbauer/Writing_a_reproducable_paper_with_quarto/).'
+You can follow any updates on the [github repository](https://github.com/paulcbauer/Writing_a_reproducable_paper_with_quarto/).
+To compile this PDF document please first read @sec-prerequisites.'
 bibliography: references.bib
@@
 
-# Prerequisites {#sec-prerequisites}
+# Prerequisites and installation {#sec-prerequisites}
 
@@
 
-To name prerequisites, we assume that you are familiar with using R and R Markdown.
+We assume that you are familiar with using R and R Markdown.
 If you don't know what R Markdown is there are many great resources that you can use (e.g.
 watch this [short video](https://vimeo.com/178485416)).
-An older template of ours [see @Bauer2018-dl; https://osf.io/q395s] may provide a quick entry point to writing a reproducible manuscript with R Markdown and Latex.
+An older template of ours [see @Bauer2018-dl] ([https://osf.io/q395s](https://osf.io/q395s)) may provide a quick entry point to writing a reproducible manuscript with R Markdown and Latex.
 To further deepen your understanding of the different components underlying this template, go on and learn about R, [Markdown](https://en.wikipedia.org/wiki/Markdown), [R Markdown](https://rmarkdown.rstudio.com/lesson-1.html) and [Quarto](https://quarto.org/).
@@
 
-Then...
+To be able to compile the present document you need to follow the steps below:
 
--   ...install [R](https://www.r-project.org/) and [Rstudio](https://posit.co/download/rstudio-desktop/) (important: most recent versions, e.g. RStudio v2022.07) [@R2017; @Rstudio2015].
--   ...install the "rmarkdown"-package from CRAN using the code below [@markdown2017].
+1. Install [R](https://www.r-project.org/) and [Rstudio](https://posit.co/download/rstudio-desktop/) (important: most recent versions, e.g. RStudio v2022.07) [@R2017; @Rstudio2015].
+2. Install [tinytex](https://yihui.name/tinytex/), a lightweight version of Tex Live [@markdown2017; @tinytex] in order to compile PDFs using the code below.
 
-``` {{r}}
-install.packages("rmarkdown")
-```
-
-```{r eval=F}
-install.packages("rmarkdown")
-```
-
--   ...install [tinytex](https://yihui.name/tinytex/), a lightweight version of Tex Live [@markdown2017; @tinytex] in order to compile PDFs.
-
-``` {{r}}
+```{r eval=F, echo=T}
 install.packages('tinytex')
@@
 
-```{r eval=F}
-install.packages('tinytex')
-tinytex::install_tinytex()
+3.
+Install the "rmarkdown"-package [@markdown2017] and the other packages [@knitr1; @knitr2; @knitr3; @kableextra; @modelsummary; @gt] using the code below.
+The `p_load` function from the `pacman` package checks tries to load a package, if not installs and loads it.
+
+```{r eval=T, echo=T}
+# Check if pacman is installed if not install
+if (!require("pacman")) install.packages("pacman")
+
+# Install/load packages if not present
+pacman::p_load(
+  rmarkdown,
+  knitr,
+  kableExtra,
+  modelsummary,
+  knitr,
+  gt,
+  gtsummary
+)
 ```
 
--   ...also install the packages below using the code below [@knitr1; @knitr2; @knitr3; @kableextra; @modelsummary; @gt].
+4.
+Download the 5 input files we created (see below) from [github](https://github.com/paulcbauer/Writing_a_reproducable_paper_with_quarto/) (click on ![](githubcode.png) $\Rightarrow$ **Download ZIP**).
+Ignore/delete the other files.
+-   `paper.qmd`: The underlying quarto markdown file.
+-   `references.bib`: The bibliography file (*We use paperpile to manage references and export the `.bib` file into the folder that contains the `.qmd` file*)
+-   `american-sociological-association.csl`: The file storing the bibliography style.
+-   `data.csv`: Some raw data.
+-   `githubcode.png`: The image file used in the abstract.
 
-``` {{r}}
-install.packages(c("knitr", "kableExtra", "modelsummary", "gt", "gtsummary"))
-```
-
-```{r eval=F}
-install.packages(c("knitr", "kableExtra",
-                   "modelsummary", "knitr", "gt", "gtsummary"))
-```
-
-- ...download the 4 input files we created (see below) from [github](https://github.com/paulcbauer/Writing_a_reproducable_paper_with_quarto/) (click on ![](githubcode.png) $\Rightarrow$ **Download ZIP**).
-Ignore/delete the other files.
-
-		- `paper.qmd`
-		- `references.bib`
-		- `american-sociological-association.csl`
-		- `data.csv`
-
-- ...finally store all files from above together in one folder (and use this folder as your working directory later on).
-
-# Basics: Input files
-
-All the input files you need to produce the present PDF file are:
-
--   `paper.qmd` (the underlying R Markdown file).
--   `references.bib` (the bibliography).
-    -   We use paperpile to manage references and export the `.bib` file into the folder that contains the `.qmd` file.
-- `american-sociological-association.csl` (the bibliography style)
--   `data.csv` (some raw data).
-
-[Download these files from github](https://github.com/paulcbauer/Writing_a_reproducable_paper_with_quarto/) (click on ![](githubcode.png) $\Rightarrow$ **Download ZIP**) and save them into a folder.
-Close R/Rstudio and directly open `paper.qmd` with RStudio.
-Doing so assures that the working directory is set to the folder that contains `paper.qmd` and the other files.[^1]
+5.
+Finally store these four files together in one folder (and use this folder as your working directory later on), close any open Rstudio instances and directly open `paper.qmd` with a new Rstudio instance.
+Doing so assures that the working directory is set to the folder that contains `paper.qmd` and the other files.
+Then click on "Render".[^1]
 
@@
 
-# Basics: Output files and the YAML header {#sec-basics-2}
+# Output files and the YAML header {#sec-basics-2}
 
@@
 
-# Compiling the document {#sec-compiling}
+# Compiling/rendering the document {#sec-compiling}
 

### 2023-04-07 paper.qmd
Small correction

@@
 5.
-Finally store these four files together in one folder (and use this folder as your working directory later on), close any open Rstudio instances and directly open `paper.qmd` with a new Rstudio instance.
+Finally store these files together in one folder (and use this folder as your working directory later on), close any open Rstudio instances and directly open `paper.qmd` with a new Rstudio instance.
 Doing so assures that the working directory is set to the folder that contains `paper.qmd` and the other files.

### 2023-04-17 paper.qmd
HTML output in YAML missing

@@
 
--   **Reproducability**: Even if you have written a study and analyzed the data yourself you will forget what you did after a few months.
+-   **Reproducibility**: Even if you have written a study and analyzed the data yourself you will forget what you did after a few months.
 A fully reproducible setup will help you to trace back your own steps.
@@
 Install the "rmarkdown"-package [@markdown2017] and the other packages [@knitr1; @knitr2; @knitr3; @kableextra; @modelsummary; @gt] using the code below.
-The `p_load` function from the `pacman` package checks tries to load a package, if not installs and loads it.
+The `p_load` function from the `pacman` package tries to load a package, if not installs and loads it.
 
@@
   rmarkdown,
-  knitr,
   kableExtra,
@@
 
-If you want your document to be rendered in multiple output formats (e.g., HTML and PDF like we did here), you need to specify both outputs in your YAML header.
-Take `paper.qmd` (the underlying quarto file of this pdf) and have a look at the YAML (line #14 - #16) to see how to specify different outputs.
+If you want your document to be rendered in multiple output formats (e.g., HTML and PDF, like we did here), you need to specify both outputs in your YAML header.
+**Take `paper.qmd` (the underlying quarto file of this pdf) and have a look at the YAML (line #14 - #16) to see how to specify different outputs.** (There is just pdf output specified no html)
 
@@
 
-To see how referencing works simply see the different examples for figures, tables and sections below.
-For instance in @sec-tables you can find different ways of referencing tables.
+To see how referencing works, simply see the different examples for figures, tables and sections below.
+For instance, in @sec-tables you can find different ways of referencing tables.
 `paper.qmd` (the underlying quarto file of this PDF) will show you how we referenced @sec-tables right here namely with '`@sec-tables`' whereas the corresponding section title was assigned the corresponding label '`# Tables {#sec-tables}`'.
@@
 
-I would recommend to use the command below and simply add it to the appendix as we did here in Appendix @sec-rsessioninfo.
+We would recommend to use the command below and simply add it to the appendix as we did here in Appendix @sec-rsessioninfo.
 This will make sure you always inform the reader about the package versions your relied on in your paper.
@@
 ``` {{r}}
-|echo=T
-|results="raw"
+#| echo=T
+#| results="raw"
 data <- read.csv("data.csv")
@@
 So you could paste that code into your `.qmd` file if you don't want to have extra data files.
-This makes sense were data files are small.
+This makes sense where data files are small.
 
@@
 For instance, imagine a reviewer asks you to omit certain observations from your sample.
-You can simply do so in the beginning of your code and push play subsequently..
-at time you might have to set `cache = FALSE` at the beginning so that all the code chunks are rerun.\
+You can simply do so in the beginning of your code and push play subsequently.
+At time you might have to set `cache = FALSE` at the beginning so that all the code chunks are rerun.\
 Researchers often avoid referring to results in-text etc.
@@
 
-Below in @fig-2 we also show a example witrh R's `ggplot2` package.
+Below in @fig-2 we also show a example with R's `ggplot2` package.
 
@@
 If you click it without using the drop down menu, by default it will create HTML since it is the first format listed in our YAML.
-Always remember to specify your desired output formats in your YAML (as described in @sec-basics-2) If you use the drop down menu of the "Render"-button, you can specify the desired output (e.g., HTML or PDF) but only if you specified it in your YAML.
+Always remember to specify your desired output formats in your YAML (as described in @sec-basics-2) **If you use the drop down menu of the "Render"-button, you can specify the desired output (e.g., HTML or PDF) but only if you specified it in your YAML.** (again: not possible here because just one output defined)
 
@@
 
-Here you can also again specify the name of the resulting document as well as the desired output format(s):
+Here you can again specify the name of the resulting document as well as the desired output format(s):
 
@@
     -   Use an one of the online services to count your words (search for "pdf word count")
--   Appendix: You can change the numbering format for the appendix in the rmd file
+-   Appendix: You can change the numbering format for the appendix in the `.qmd` file
     -   What is still not possible in this document is to automatically have separate reference sections for paper and appendix.

### 2023-04-18 paper.qmd
fixed YAML header with HTML output, changed two reference to this in the text

@@
 format:
+  html:
+    toc: true
+    toc_float: true
+    toc_depth: 2
+    number_sections: true
+    code_folding: hide
+    theme: readable
   pdf:
@@
 If you want your document to be rendered in multiple output formats (e.g., HTML and PDF, like we did here), you need to specify both outputs in your YAML header.
-**Take `paper.qmd` (the underlying quarto file of this pdf) and have a look at the YAML (line #14 - #16) to see how to specify different outputs.** (There is just pdf output specified no html)
+Take `paper.qmd` (the underlying quarto file of this pdf) and have a look at the YAML (line #15 - #22) to see how to specify different outputs.
 
@@
 If you click it without using the drop down menu, by default it will create HTML since it is the first format listed in our YAML.
-Always remember to specify your desired output formats in your YAML (as described in @sec-basics-2) **If you use the drop down menu of the "Render"-button, you can specify the desired output (e.g., HTML or PDF) but only if you specified it in your YAML.** (again: not possible here because just one output defined)
+Always remember to specify your desired output formats in your YAML (as described in @sec-basics-2) If you use the drop down menu of the "Render"-button, you can specify the desired output (e.g., HTML or PDF) but only if you specified it in your YAML.
 

### 2023-04-25 paper.qmd
Update to highlight compilation issue.

@@
 You can follow any updates on the [github repository](https://github.com/paulcbauer/Writing_a_reproducable_paper_with_quarto/).
-To compile this PDF document please first read @sec-prerequisites.'
+To compile this PDF document please first read @sec-prerequisites (known issues: if folder is part of a cloud drive, html compilation sometimes fails).'
 bibliography: references.bib
